\section{Extensions and Conclusion}
\label{sec:limitations}

The theorem deliberately factors correctness failures from the quantitative
noise-flooding argument.  Its support-level hypotheses prove the good-execution
case.  For an imperfectly correct instantiation, let
\(\mathsf{Bad}_{\mathrm{corr}}\) be the event that any key-generation,
encryption, or evaluation result used by the execution violates the current
support invariant.  One standard up-to-bad transition from the real game to
the good-execution game then gives
\[
  \Pr[\mathsf{IND\text{-}CPAD}=1]
  \leq \beta + \frac{\sqrt{qn}}{2\gamma}
      + \Pr[\mathsf{Bad}_{\mathrm{corr}}].
\]
Thus the imperfect-correctness result is one modular game hop away from
\Cref{thm:main}, not a redesign of its Pythagorean argument.  What remains for
a concrete instantiation is an aggregate bound on
\(\Pr[\mathsf{Bad}_{\mathrm{corr}}]\).  Such a bound may be supplied directly;
if it is derived from per-operation failure probabilities, the surrounding
security experiment must also account for encryption and evaluation calls.

The local charts in \Cref{eq:charts} need a concrete construction and
proof for the intended plaintext representation, including wraparound and
decoding behavior.  A CKKS-level result also requires connecting the abstract
scheme operations and stored error field to an implementation and parameter
analysis.  As in SSProve's own computational-security interpretations, the
probabilistic-polynomial-time condition is a conventional metatheoretic audit,
because the package semantics do not internalize a runtime model~\cite{ssprove}.
That audit is routine here: the reduction invokes the adversary once, forwards
its encryption and evaluation calls, answers at most \(q\) selected decryption
calls, and performs table bookkeeping and Gaussian sampling.  Formal automatic
complexity analysis would be useful, but is orthogonal to this work.  Finally,
the released SSProve/MathComp dependency graph still names an admitted
real-summation lemma.  The artifact supplies a kernel-checked proof of that
exact statement, and the proof has been submitted upstream.  No further local
mathematical obligation remains; a future dependency update can remove the
name from the top-level assumption report without changing the security proof.

Within this boundary, the artifact checks the central security argument: a
one-call discrete-Gaussian KL bound can be lifted through adaptive, stateful
decryption queries without collapsing to a linear hybrid loss.  The resulting
IND-CPAD-to-IND-CPA reduction is stated over concrete SSProve games, with one
quantitative hop and explicit assumptions.  This provides a reusable route for
other cryptographic proofs whose useful parameters depend on paying divergence
once at the end of an adaptive transcript.

\section*{Ethics Considerations}

This work studies and verifies a published defense against an already public
decryption attack.  It performs no experiments on deployed systems, users, or
private data and introduces no new exploit.  Its principal risk is
overconfidence: treating the conditional theorem as a verified concrete CKKS
deployment could encourage unsafe parameters.  We mitigate that risk by
stating the good-execution boundary, required correctness-failure term, chart
and instantiation obligations, semantic model, and trusted base explicitly.
