\section{Formalization Details}
\label{sec:independent-verification}

A formal theorem is useful only if its statement is the theorem one intended
to prove.
For cryptographic verification, this creates two different review tasks.
First, one should audit the \emph{specification layer}: the games, reduction,
scheme interface, metric assumptions, and final probability bound.
Second, one may audit the proof infrastructure that connects those definitions
to the theorem checked by Rocq.
This section explains both tasks, with the first three subsections written for
a cryptographer who wants to understand the formal artifact without first
learning the details of Rocq proof engineering or program-logical soundness.

The development has three technical layers.
The probability layer proves the KL, Pinsker, Pythagorean preservation, and
discrete-Gaussian facts used in the noise-flooding calculation.
The program-logic layer packages those facts as sound reasoning principles for
SSProve games.
The LMSS/noise-flooding layer defines the games, reduction, metric interface,
and final security theorem.
The key audit question is whether the top layer says the right cryptographic
thing; the lower layers should then be read as checked proof infrastructure
for that statement.

\subsection{Artifact Accounting}
\label{sec:artifact-accounting}

\Cref{tab:artifact-accounting} is a map of the current Rocq artifact.
The counts are source lines in the named files, rounded only by the grouping
choices in the table.
They are meant to answer a practical question: where does one look for the
mathematics, where does one look for the program-logic infrastructure, and
where does one look for the cryptographic statement?

The line counts should not be read as a measure of mathematical complexity or
human effort.
Mechanized proofs contain many lines that correspond to routine bookkeeping:
transporting invariants through syntax, discharging type and support
conditions, or restating a familiar argument in the library's preferred form.
This development was also substantially AI-assisted.
As a result, the code is less proof-golfed than a fully hand-written artifact,
and it carries more local technical debt than a fully hand-written development
would usually tolerate before refactoring.
At the same time, AI assistance changes development speed enough that
person-month estimates would be especially misleading.
We therefore report line counts as orientation data, not as a claim about
proof elegance or implementation effort.

\begin{table}
\centering
\resizebox{\linewidth}{!}{%
\begin{tabular}{lrl}
\hline
Layer & Lines & Representative files \\
\hline
KL, Pinsker, Pythagorean probability, discrete Gaussians
  & 5527
  & \texttt{Probability/KL/}, \texttt{Probability/DiscreteGaussians/} \\
Program logics and trace infrastructure
  & 6657
  & \texttt{ProgramLogics/}, \texttt{Probability/\{DletTuple,OutputHeap,PythSeq\}.v}, \texttt{NextMessage/Trace.v} \\
LMSS/noise-flooding games and security proof
  & 11199
  & \texttt{Schemes/\{ApproxFHE,Indcpa,Indcpad\}.v}, \texttt{Constructions/NoiseFlooding.v}, \texttt{Security/IndcpadSimulator.v}, \texttt{NoiseFloodingSecurity/} \\
\hline
Three main technical layers & 23383 & \\
Full Rocq development & 27938 & includes local MathComp/SSProve adapters \\
\hline
\end{tabular}
}
\caption{Coarse source-line map of the formal artifact.  These counts are
orientation data rather than effort metrics.}
\label{tab:artifact-accounting}
\end{table}

\subsection{Trusted Computing Base}
\label{sec:tcb}

The trusted computing base is the part that is accepted rather than proved in
this paper.
For this artifact, it is the usual base for a Rocq/SSProve development plus
the libraries needed to formalize countable probability.
At the bottom is the Rocq kernel and standard library infrastructure,
including the construction of real numbers and the classical reasoning
principles used by imported libraries.
MathComp and MathComp Analysis provide algebra, finite tuples, real analysis,
infinite sums, summability, logarithms, exponentials, and the distribution
interface used for KL divergence and discrete Gaussians.
SSProve provides the meaning of the pseudocode used for games and reductions:
probabilistic programs, stateful oracle packages, oracle linking, heaps,
subdistributions, and completed games.

Our claim starts above that foundation.
Given SSProve's semantics for games and MathComp's semantics for distributions
and real analysis, the formal artifact checks that the noise-flooding theorem
follows from the stated scheme, correctness, metric, and IND-CPA assumptions.
The proof-specific ingredients are not additional trusted axioms.
The program-logic rules, trace compiler, KL lemmas, discrete-Gaussian facts,
and individual game hops are Rocq theorems proved from the underlying
semantics.
Thus a crypto reader need not trust a new informal program logic on faith:
the logic is a proof organization device whose rules are themselves checked
theorems.

If a reader does not accept Rocq, MathComp/Analysis, or SSProve as a
foundation, this paper cannot remove that concern; many textbooks' worth of
mathematics and programming-language semantics live below our development.
Once those foundations and the cryptographic definitions are accepted, the
remaining machine-checkable claim is deliberately simple: Rocq typechecks the
proof.
The command
\[
  \texttt{Print Assumptions Secure.is\_secure}
\]
is the machine-checkable summary of what remains outside the proof: the
scheme/correctness/metric interfaces, the IND-CPA game interface, positivity
of the Gaussian-width multiplier, and the ordinary library/classical
assumptions inherited from Rocq, MathComp, MathComp Analysis, and SSProve.

\subsection{What a Cryptographer Should Audit}
\label{sec:crypto-audit-surface}

The main theorem is exposed as
\texttt{NoiseFloodingSecure.is\_secure}.
The statement itself is the primary cryptographic object to audit.
It says that for every adversary accepted by the IND-CPAD interface, the
noise-flooded winning probability is at most the IND-CPA winning probability
of a concrete constructed reduction plus the verified noise-flooding loss.
The word ``package'' in the formal statement should be read simply as
``stateful oracle program''; the important cryptographic point is that the
adversary can make adaptive encryption, evaluation, and decryption queries
subject to the stated decryption-query bound.

A crypto-facing audit can therefore focus on four questions.
Does the IND-CPAD game encode the intended Li--Micciancio-style decryption
attack model?
Does the IND-CPA game on the right-hand side match the assumed security notion
for the underlying scheme?
Is the concrete reduction the expected one, with simulated decrypt queries
answered by noise flooding the common table plaintext?
Do the metric, correctness, and Gaussian-width assumptions justify the loss
\(2\sqrt{q\epsilon_{\mathsf{nf}}/2}\), where
\(\epsilon_{\mathsf{nf}}=n/(2\gamma^2)\)?
The appendix excerpts are organized around exactly these questions.
\Cref{app:ind-cpa-reduction-excerpt} shows the concrete reduction package;
\Cref{app:security-bound-excerpt} shows the reduction-shaped bound as the
definition of \texttt{security\_bound}; \Cref{app:reduction-bound-excerpt}
shows that the constructed reduction has the IND-CPA adversary interface; and
\Cref{app:final-reduction-theorem} shows the final theorem statement.

The crypto-facing audit surface is small compared with the proof body.
\Cref{tab:crypto-audit-surface} lists the files whose definitions and
assumptions determine what the theorem means.
The line counts are the excerpted appendix lines, included to show the size of
the human-facing audit surface rather than the size of the proof scripts.

\begin{table}
\centering
\resizebox{\linewidth}{!}{%
\begin{tabular}{lrl}
\hline
File & Excerpt lines & What to audit \\
\hline
\texttt{Schemes/ApproxFHE.v} & 83
  & scheme interface, correctness assumptions, metric charts \\
\texttt{Schemes/Indcpa.v} & 60
  & IND-CPA game and adversary-dependent bound interface \\
\texttt{Schemes/Indcpad.v} & 108
  & IND-CPAD game, oracle table, decrypt-query behavior \\
\texttt{Constructions/NoiseFlooding.v} & 45
  & transformed scheme and Gaussian-width parameter \\
\texttt{Security/IndcpadSimulator.v} & 95
  & concrete IND-CPA reduction and simulation oracle \\
\texttt{NoiseFloodingSecurity/Prelude.v} & 5
  & closed-form noise-flooding loss \\
\texttt{NoiseFloodingSecurity/Final.v} & 13
  & security bound, reduction validity, main theorem statement \\
\hline
Total & 409 & crypto-facing specification layer \\
\hline
\end{tabular}
}
\caption{Primary cryptographic audit surface.  The files are under
\texttt{theories/}; \Cref{app:formal-interface-excerpts} includes exactly
these excerpts, while the rest of the development proves consequences of these
definitions and assumptions.}
\label{tab:crypto-audit-surface}
\end{table}

The appendix therefore includes selected excerpts from the specification layer,
not the whole proof script.
This is deliberate: the specification layer is where a cryptographer checks
that the theorem is about the intended games and assumptions.
\Cref{app:approximate-homomorphic-encryption-interface,app:metric-chart-excerpt,app:support-correctness-excerpt}
show the scheme interface and correctness/metric assumptions;
\Cref{app:ind-cpa-game-excerpt,app:ind-cpa-d-game-excerpt} show the two games;
\Cref{app:noise-flooded-construction-excerpt,app:ind-cpa-reduction-excerpt,app:loss-formula-excerpt,app:security-bound-excerpt}
show the construction, concrete reduction, closed-form loss, and
reduction-shaped security bound.
\Cref{app:reduction-bound-excerpt} shows reduction validity, and
\Cref{app:final-reduction-theorem} shows only the main theorem statement.
The proof scripts are not the part a cryptographer should have to read in
order to understand what has been proved.
The full development is about 28,000 lines, with just under 10,000 lines in the
split noise-flooding security proof alone.
The artifact is the appropriate medium for that part.

The assumptions that should receive the most cryptographic attention are the
following.
The IND-CPA interface in \texttt{Indcpa.v} gives the game used on the
right-hand side of the reduction theorem.
One may separately instantiate it with a concrete computational security
assumption for the underlying scheme.
The reduction in \texttt{IndcpadSimulator.v} is also on the hot path.
The formal theorem proves a statement about the concrete SSProve package in
\Cref{app:ind-cpa-reduction-excerpt}; it is not an existential reduction.
A computational interpretation still requires the usual runtime audit: the
reduction should run in time comparable to the original adversary, plus the
explicit bookkeeping for oracle tables, homomorphic evaluation, and Gaussian
sampling.
The Rocq development does not formalize a polynomial-time cost model, so this
runtime check remains a conventional cryptographic review obligation.
The IND-CPAD game in \texttt{Indcpad.v} is not assumed secure; it is the game
proved secure after applying noise flooding.
The final theorem currently uses the support-level perfect-correctness
interface in \texttt{ApproxFHE.v}: every ciphertext row that the challenger can
store decrypts to the deterministic center required by the metric argument.
Approximate correctness with explicit correctness-error accounting is kept as
a separate interface.
The metric interface in \texttt{ApproxFHE.v} assumes origin-centered integer
charts:
\[
  I_a(a)=0,\qquad
  d(a,b)=\lVert I_a(b)\rVert,\qquad
  I_b^{-1}(x)=I_a^{-1}(x+I_a(b)).
\]
These are the assumptions connecting the abstract plaintext metric to the
integer-vector Gaussian analysis in \Cref{sec:isometry}.
Finally, the noise-flooding parameter module assumes only that the Gaussian
width multiplier is positive; the quantitative dependence on that multiplier
is the explicit loss formula above.

\subsection{Probability and Discrete-Gaussian Review}
\label{sec:probability-review}

The probability layer is the part of the artifact that a mathematically minded
reader should audit as ordinary analysis.
The relevant human-facing statements are the definitions of KL divergence and
finite KL, Pinsker's inequality for MathComp distributions, the conditional
coordinate chain bound used for Pythagorean preservation, and the
equal-variance KL formula for discrete Gaussians over the integers.
These are stated in \Cref{sec:verified-kl-analysis}; their Rocq proofs live
primarily in \texttt{Probability/KL/} and
\texttt{Probability/DiscreteGaussians/}.

As with the program logic, the full proof scripts are not the intended reading
order.
The paper should give pen-and-paper proofs of the mathematical facts that are
not immediate, especially the chain-rule/summability argument and the
discrete-Gaussian KL calculation.
The formal development then checks the details that are easy to lose in an
informal proof over countable distributions: absolute continuity,
positive/negative-part summability, interchange of countable sums, and the
normalization and moment facts for the integer Gaussian.

This part of the project also produced some library-adjacent ``community
service'' for MathComp Analysis.
The local \texttt{RealSumExtras.v} file factors out reusable facts about
MathComp's experimental \texttt{sum}/\texttt{summable} interface, including the
proved Fubini-style interchange lemma
\[
  \texttt{interchange\_psum\_proved}.
\]
Those facts were developed while discharging the summability obligations for
the KL analysis.
A version of this material was also submitted upstream to MathComp Analysis,
where it helps reduce reliance on admitted real-summation infrastructure.

\subsection{Program-Logic Review}
\label{sec:program-logic-review}

For a PL or verification reader, the audit story is the standard soundness
story.
The development defines semantic judgments for support-style Hoare reasoning,
completed-output additive error, and Pythagorean KL bookkeeping.
Each inference rule is a theorem about those semantic definitions.
A human reader should not be expected to step through the corresponding Rocq
proof scripts line by line.
The human-facing artifact is the selected collection of semantic definitions
and rule statements, together with mathematical explanations for the few rules
whose soundness is not immediate.

The rule statements themselves are compact.
\Cref{sec:program-logic} displays the core rules used by the security proof:
completed additive-error rules, Pythagorean sampling and sequencing rules, the
Micciancio--Walter bridge, and the trace-compiler replacement principles.
Even conservatively counting around twenty externally meaningful rules at ten
to twenty lines each, the surface to inspect is on the order of a few hundred
lines, not thousands of lines of proof script.

The proof scripts remain important, but as the machine-checking substrate.
For routine rules, soundness is a direct unfolding of the semantic judgment
plus standard distribution facts.
For the nontrivial rules---completed-output sequencing, the trace compiler,
and the Pythagorean final-marginal theorem behind the Micciancio--Walter
bridge---the paper should provide pen-and-paper arguments at the level of
mathematics and proof structure.
The Rocq proofs then check that those arguments have no hidden side conditions.

For cryptographers, this part should look familiar at a high level.
We define judgments by their intended semantics, prove inference rules sound,
and then use those rules to structure the game-hopping proof.
The unusual feature is not the soundness methodology itself, but the
Pythagorean bookkeeping: instead of composing statistical-distance errors
linearly at each program step, the proof records conditional KL costs and pays
the square-root loss once at the end.
