\section{Introduction}

Fully homomorphic encryption (FHE) lets a server compute on encrypted data
without decrypting it.
Approximate schemes such as CKKS \cite{ckks} make this idea practical for
real-valued arithmetic, but they also make security statements more delicate:
decryption returns an approximate value, and the approximation error itself can
become a source of information.
The Li--Micciancio attack \cite{lm} showed that this issue is not merely
cosmetic.
Later work has shown that this decryption-query viewpoint is broader than
CKKS: exact FHE schemes can also fail IND-CPAD when parameters are pushed so
aggressively that decryption errors become visible \cite{cheon2024attacks}.
For exact schemes, the natural response may simply be to restore exact
correctness by choosing less aggressive parameters.
For approximate FHE, however, approximate decryption is part of the interface,
so the security question is how to hide the approximation error while keeping
the answer useful.
Li, Micciancio, Schultz-Wu, and Sorrell \cite{lmss} proposed a noise-flooding
defense and proved that, under suitable parameters, the resulting scheme
satisfies a stronger decryption-query security notion.

This proof is exactly the kind of cryptographic argument that should benefit
from mechanization.
It combines game-hopping, oracle reductions, approximate correctness,
discrete-Gaussian calculations, and a sharp accounting of many related random
samples.
The last point is important.
A naive hybrid argument would pay linearly for $q$ decryption queries.
The cryptographic analysis instead uses the Micciancio--Walter
``Pythagorean'' phenomenon \cite{mw17}: when each step has a bounded
conditional KL divergence, the final statistical distance is controlled by the
square root of the sum of the KL costs.
This is the difference between a linear loss and the expected
$O(\sqrt q)$ noise-flooding loss.

We formalize this proof in Rocq using SSProve \cite{ssprove}.
Our theorem is deliberately stated for an abstract approximate-FHE interface
rather than for a concrete CKKS implementation.
The interface supplies the IND-CPA assumption for the underlying scheme, a
support-level correctness assumption, and origin-centered integer-vector
metric charts that connect plaintext distance to discrete-Gaussian sampling.
For every adversary in the IND-CPAD game, the formal development constructs
the usual IND-CPA reduction and proves the bound
\[
  \Pr[\mathsf{IND\text{-}CPAD}_{\mathsf{NF}}(\cA)=1]
  \leq
  \Pr[\mathsf{IND\text{-}CPA}(\mathcal{B}_{\cA})=1]
  +
  \frac{\sqrt{qn}}{\gamma}.
\]
Here $q$ is the decryption-query bound, $n$ is the metric-chart dimension, and
$\gamma$ is the Gaussian-width multiplier.
The displayed loss should be read as a parameter tradeoff, not only as a
theorem about distinguishing probabilities.
The \(q\)-dependence can be cancelled by increasing \(\gamma\) like
\(\sqrt q\), up to the dimension factor.
In the construction summarized below, \(\gamma\) scales linearly with the
standard deviation of the flooding noise, after the flood has already been
scaled to dominate the allowed correctness error.
Thus increasing \(\gamma\) is not a harmless proof maneuver: it directly means
adding wider noise to decryption answers.

For intuition, take \(q=2^{64}\) decryption queries and ignore constants and
the fixed dimension factor.
The square-root bound asks the flooding width to absorb a factor
\(\sqrt q=2^{32}\), while a linear hybrid bound would ask it to absorb
\(q=2^{64}\).
Thinking of standard deviation as the relevant width, this is roughly the
difference between losing 32 and 64 extra bits of precision to the flooding
noise.
At the extreme, one can make the scheme trivially secure by choosing
\(\gamma\) so large that the returned approximate plaintext is drowned in
fresh noise; the message is then effectively masked by the flood.
Such a defense is useful only if the answer remains accurate enough for the
application.
Recovering accuracy after adding more noise pushes cost elsewhere, for
example into larger parameters, more noise-budget headroom, or additional
bootstrapping.
Since these resources are a major source of FHE's runtime cost, avoiding an
unnecessary \(O(q)\) loss is part of making the security argument compatible
with usable approximate homomorphic encryption.

This parameter-choice problem is a subject in its own right.
Recent work has studied precision loss in approximate homomorphic encryption
and revisited noise-flooding countermeasures from the point of view of
concrete parameter selection
\cite{costache2023precision,bergamaschi2025revisiting}.
Our theorem verifies one sufficient analytic way to choose the flooding width
and obtain a provable IND-CPAD reduction; it is not meant to be the last word
on optimizing this tradeoff.
The theorem is exposed in the artifact as
\texttt{NoiseFloodingSecure.is\_secure}.

\subsection{Why a Program Logic?}

The central proof-engineering problem is preserving this
$O(\sqrt q)$ bound inside the mechanized game-hopping proof.
Standard approximate relational logics are built around sequencing rules that
add the error of the first command to the error of the continuation.
This is exactly right for ordinary statistical-distance or failure-probability
arguments, but it is too weak for noise flooding: composing $q$ decryption
replacements with such a rule would pay at least a linear $O(q)$ loss.
The cryptographic proof gets a sharper result because it does not pay
statistical distance at each call.
Instead, it records a KL cost for each conditional sampling step and converts
the whole tuple of costs to statistical distance only once, using the
Pythagorean preservation theorem.

Our program logic is designed to expose that accounting discipline at the
level of SSProve syntax.
The additive-error judgment is the ordinary cryptographic judgment used for
game hops.
The Pythagorean judgment is the nonstandard part: its sequencing rules
concatenate KL-budget tuples instead of immediately adding statistical-distance
errors, and the Micciancio--Walter bridge later turns a budget $s$ into the
single loss $\sqrt{\lVert s\rVert_1/2}$.
Thus the logic is what makes the formal proof express the same
``pay once at the end'' idea as the paper cryptographic argument.

\subsection{Paper Organization}

\Cref{sec:prelim-probability-notation} fixes notation for distributions,
statistical distance, KL divergence, SSProve programs, metric charts, and the
noise-flooding parameters.
\Cref{sec:verified-kl-analysis} gives the verified probability facts used by
the rest of the paper: Pinsker's inequality, the conditional-coordinate
Pythagorean preservation theorem, and the KL calculation for equal-variance
integer Gaussians.
\Cref{sec:program-logic} then explains the program-logic layer that lets those
probability facts be applied to SSProve games and adaptive oracle code.
\Cref{sec:noise-flooding-proof} is the cryptographic application: it states
the abstract approximate-FHE interface, describes the IND-CPAD-to-IND-CPA
hybrid chain, isolates the one nonzero noise-flooding hop, and records the
final theorem.
\Cref{sec:independent-verification} explains how to audit the formal
artifact, and the appendices reproduce the main game and interface excerpts.

The paper is meant to support more than one reading path.
Readers interested in the formal program logic should read
\Cref{sec:verified-kl-analysis,sec:program-logic} before the application.
Readers whose main interest is the cryptographic theorem can instead read
\Cref{sec:noise-flooding-proof,sec:artifact-accounting,sec:tcb,sec:crypto-audit-surface}
as a largely self-contained account of the verified LMSS-style argument and
its audit surface.
These parts are of independent interest to theory cryptographers: they are
written to minimize assumed programming-language or formal-verification
prerequisites, while assuming standard cryptographic familiarity with games,
reductions, hybrids, and oracle adversaries.

\subsection{Contributions}

\paragraph{Verified probability analysis.}
We formalize the probability facts needed by the program logic over MathComp
distributions, including KL divergence for countable distributions, absolute
continuity and finite-KL side conditions, Pinsker's inequality, and a
conditional-coordinate chain theorem that realizes Pythagorean probability
preservation.
We also prove the equal-variance KL formula for discrete Gaussians over the
integers,
\[
  \KL{\DG{\mu}{\sigma^2}}{\DG{\nu}{\sigma^2}}
  =
  \frac{(\nu-\mu)^2}{2\sigma^2},
\]
including the summability and moment facts needed by the formal proof.
Some reusable real-summation lemmas developed for this part were also
submitted upstream to MathComp Analysis.

\paragraph{Relational reasoning for Pythagorean errors.}
We define additive-error and Pythagorean relational judgments directly over
SSProve semantics.
The judgments use completed output distributions, so missing subdistribution
mass is represented by an explicit failure point instead of an informal side
condition.
We prove the core structural, sampling, sequencing, and
Micciancio--Walter bridge rules in Rocq.
The rules are semantic theorems, not axioms.

\paragraph{Trace compilation for oracle replacement.}
SSProve adversaries are concrete programs, not opaque modules.
We exploit this by verifying a trace compiler that exposes the first $q$ calls
to a selected oracle and records the residual computation as replayable traces.
The compiler puts adaptive adversaries into the fixed-call-schedule form
expected by the Pythagorean logic.
The main compiler theorem then lifts a one-call Pythagorean replacement proof to a
whole-adversary additive-error bound of
$\sqrt{q\lVert s\rVert_1/2}$.
This is the mechanism that turns the one-decryption Gaussian calculation into
the $q$-query noise-flooding bound.

\paragraph{Verified LMSS-style noise flooding.}
Combining the probability library, program logic, and trace compiler, we prove
the noise-flooding security theorem for an abstract approximate-FHE scheme.
The final statement is an inequality about SSProve winning probabilities, not
about derivability in our logic.
Thus the program logic organizes the proof, while the exported theorem remains
anchored in the standard SSProve game semantics.
The accompanying Rocq development is available at
\url{https://github.com/ethanlee515/mending}.
