\section{Noise Flooding for Approximate Homomorphic Encryption}
\label{sec:noise-flooding-proof}

This section gives the cryptographic proof carried by the formal development.
It is written so that the main argument can be read as an ordinary
game-hopping proof: the verification-specific machinery is used to justify
the hops, but the objects being related are the usual security games,
oracles, reductions, and probability bounds.

The theorem is parameterized by an abstract approximate homomorphic encryption
scheme.
It is intentionally not a CKKS instantiation.
The scheme interface, correctness assumptions, IND-CPA assumption, and metric
assumptions are all explicit parameters, and the final formal theorem is
\texttt{NoiseFloodingSecure.is\_secure}.
Thus the statement should be read as: any concrete approximate-FHE scheme that
implements this interface inherits the verified noise-flooding reduction.

The security notion is the IND-CPAD game for Li--Micciancio-style decryption
attacks on approximate FHE.
The challenger and oracle definitions are given in \Cref{app:ind-cpad-game};
the main body uses only their standard roles.
The adversary receives encryption, evaluation, and decryption oracles, and the
decryption oracle enforces a bound $q$ by maintaining a counter.
One piece of terminology is specific to SSProve but harmless: a
\emph{package} is a stateful implementation of one or more oracles.
The adversary is otherwise an arbitrary oracle program; in particular, it is
not assumed to have been written in advance as a list of $q$ separate
decryption-query routines.

\subsection{Metric and Correctness Interface}
\label{sec:isometry}

Noise flooding compares two plaintext centers by sampling discrete Gaussian
noise around them.
To state that comparison without committing to a concrete message space, the
proof assumes a message type $M$ with a metric $d : M \times M \to \bbN$ and
local integer-vector charts.
For each center $a \in M$, the chart
\[
  I_a : M \to \bbZ^n
\]
represents nearby plaintexts as integer vectors relative to $a$, with an
inverse chart map taking vectors back to messages.
The interface used by the current theorem is origin-centered:
\[
  I_a(a)=0,
  \qquad
  d(a,b)=\|I_a(b)\|,
\]
together with the chart-compatibility equation
\[
  I_b^{-1}(x)=I_a^{-1}(x+I_a(b)).
\]
The first two equations say that the center has vector representative zero and
that the message metric is exactly the vector norm in the chart.
The compatibility equation says that changing the center from $a$ to $b$
amounts to translating coordinates by the vector for $b$ in the $a$-chart.
This abstracts the usual modular-representative bookkeeping in
lattice-based approximate FHE: once a plaintext is chosen as the local center,
nearby plaintexts are represented by short integer vectors, and translating
the center translates the coordinate system.

The correctness assumption is also stated at the interface level.
The proof uses the support-level interface
\texttt{ApproxCorrectnessPerfect}.
The scheme decryption algorithm remains distribution-valued,
but the correctness and metric statements use a deterministic decryption center
\[
  \mathsf{deterministic\_dec}(sk,c).
\]
The assumption \texttt{deterministic\_dec\_correct} says that the actual
decryption distribution is the point mass at this center.
Encryption and evaluation correctness then give the property needed by the
security proof: every ciphertext row that can appear in the challenger's table
decrypts within its recorded error bound of the corresponding challenge
plaintext.
Equivalently, whenever the decryption oracle reaches a well-formed table row,
the real decryption center and the table plaintext are close in the metric
above.

\subsection{Query-Bounded Oracle Adversaries}
\label{sec:query-bound}

The only structural obstacle to a direct hybrid proof is adaptivity.
For a fixed list of $q$ decryption calls, the Pythagorean analysis would simply
replace one call at a time and combine the $q$ KL costs.
An oracle adversary does not present such a list: it chooses whether and when
to call decryption based on previous answers.

The trace compiler from \Cref{sec:program-logic} is the formal device that
bridges this gap.
Given an ordinary SSProve adversary and a selected operation identifier, the
compiler exposes the first $q$ calls to that operation and records the
surrounding computation as a replayable trace.
For a cryptographer, this can be read as a verified version of the familiar
``run until the next oracle call, answer it, and resume'' transformation.
Invalid replay branches are interpreted as null code, so the compiler theorem
is a semantic statement about the adversary's behavior rather than a claim
about a particular syntactic normal form.

In the noise-flooding proof, the selected operation is the decryption oracle.
The real and simulated decryption oracles both maintain the same decrypt-count
location.
After $q$ successful selected calls, any later decrypt call assert-fails on
both sides.
Thus replacing the first $q$ selected calls is enough to cover every
non-failing decryption response, even though the adversary itself may decide
when to query decryption in a data-dependent way.

\subsection{Proof Overview}
\label{sec:games}

The proof follows the standard cryptographic shape: define a sequence of
games, prove neighboring games close, and compose the losses.
The endpoints are the usual ones:
$H_0$ is the real IND-CPAD experiment for $\cA$, and the final game is the
IND-CPA experiment for the reduction adversary $\mathcal{B}_{\cA}$.
The reduction is the standard one: it answers encryption queries using the
IND-CPA challenge oracle, recomputes homomorphic-evaluation table entries
locally, and answers decrypt queries by flooding the common plaintext when the
two stored labels agree.
We spell out the pseudocode in \Cref{alg:ind-cpa-reduction}; the main proof
below is about why the transition to that conventional reduction is justified
inside the verified program logic.

The intermediate games are presentations of the same cryptographic experiment
with different oracle implementations and, in two places, with the trace
compiler inserted or removed.
They are useful because they isolate the only probabilistic loss from the
bookkeeping equalities.
To name the intermediate games, we use the trace compiler from
\Cref{sec:program-logic} directly.
Let $C_{\cA}^{\mathsf{open}}$ be the open IND-CPAD experiment body: it
initializes the challenger state, runs $\cA$, and returns the final guess
predicate, but its oracle calls are not yet linked to an oracle package.
Let $P_{\mathsf{real}}$ be the ordinary IND-CPAD oracle package and
$P_{\mathsf{sim}}$ be the same package with the decrypt oracle replaced by the
noise-flooding simulator.
For a package $P_s$ used by the compiler for the selected decrypt calls and a
package $P_r$ used for all calls left in the residual program, write
\[
  G(P_s,P_r)
  =
  \mathsf{Link}\bigl(
    \operatorname{compileCalls}_q(P_s,\mathsf{ODec},C_{\cA}^{\mathsf{open}}),
    P_r
  \bigr).
\]
Thus the compiler is applied to the open experiment code, not literally to
$\cA$ as a standalone adversary package.
With this notation, the hybrids are
\[
\begin{array}{rcl}
  H_0 &=& \mathsf{IND\text{-}CPAD}(\cA),\\
  H_1 &=& \mathsf{Link}(C_{\cA}^{\mathsf{open}},P_{\mathsf{real}}),\\
  H_2 &=& G(P_{\mathsf{real}},P_{\mathsf{real}}),\\
  H_3 &=& G(P_{\mathsf{sim}},P_{\mathsf{real}}),\\
  H_4 &=& G(P_{\mathsf{sim}},P_{\mathsf{sim}}),\\
  H_5 &=& \mathsf{Link}(C_{\cA}^{\mathsf{open}},P_{\mathsf{sim}}),\\
  H_6 &=& \mathsf{IND\text{-}CPA}(\mathcal{B}_{\cA}).
\end{array}
\]
$H_0$ is the security game from \Cref{app:ind-cpad-game}.
$H_1$ is the same game written as open experiment code linked with the real
oracle package.
$H_2$ is the compiled-real presentation.
The equality $H_1=H_2$ is the same-package compiler theorem:
compiling selected calls and linking them back to the same package preserves
the completed semantics.
Here ``completed'' means that assertion failure is treated as an explicit
failure output, so equalities and statistical-distance bounds compare total
output distributions.

The paper-level hybrid chain is then
\[
  H_0 \xrightarrow{0} H_1
  \xrightarrow{0} H_2
  \xrightarrow{\Delta_{\mathsf{comp}}} H_3
  \xrightarrow{0} H_4
  \xrightarrow{0} H_5
  \xrightarrow{0} H_6,
\]
where
\[
  \epsilon_{\mathsf{nf}}=\frac{n}{2\gamma^2},
  \qquad
  \Delta_{\mathsf{comp}}=\sqrt{q\epsilon_{\mathsf{nf}}/2}.
\]
Here $n$ is the chart dimension and $\gamma$ is the Gaussian-width multiplier.
Only the arrow $H_2\to H_3$ spends statistical distance.
The semantic relation between neighboring closed games is the canonical
additive-error judgment from \Cref{sec:program-logic}, interpreted over
completed outputs; the Pythagorean judgment is used only inside the proof of
the paid hop.
The Coq development refines the last exact arrow through an auxiliary linked
presentation, but this is package-linking bookkeeping rather than an additional
cryptographic endpoint.
The public theorem states the final winning-probability bound, so the exported
security loss is twice the compile distance:
\[
  \mathsf{security\_loss}(q)
    = 2\sqrt{q\epsilon_{\mathsf{nf}}/2}.
\]
This square-root dependence is parameter-critical, not just a prettier proof
bound.
Since \(\epsilon_{\mathsf{nf}}=n/(2\gamma^2)\), the displayed loss is
\(\sqrt{qn}/\gamma\).
Keeping the final loss fixed therefore requires the Gaussian-width multiplier
\(\gamma\) to grow like \(\sqrt q\) for fixed \(n\); a linear composition of
per-query statistical-distance losses would force \(\gamma\) to grow like
\(q\).
In approximate FHE, \(\gamma\) is actual added decryption noise.
Making it unnecessarily large worsens precision and noise-budget pressure, and
can force larger parameters or more bootstrapping.

\subsection{The Nonzero Hop}
\label{sec:paid-hop}

The hop $H_2 \to H_3$ is the only place where the noise-flooding analysis
pays loss.
In this hop, the compiled experiment answers each of the first $q$ selected
decrypt calls with the simulator instead of the real decrypt oracle.
Let $O_{\mathsf{real}}(i)$ and $O_{\mathsf{sim}}(i)$ be the real and simulated
decrypt oracle bodies, and let $\Phi$ be the challenger heap invariant
recording initialized keys, table validity, and the decrypt counter.

\begin{lemma}[One-call decryption replacement]
\label{lem:one-call-decryption-replacement}
For every decrypt input \(i\), the real and simulated decrypt oracle bodies
satisfy the Pythagorean judgment
\[
  \vDash\Pyth\set{i_L=i_R\wedge m_L=m_R\wedge \Phi(m_L)}
    O_{\mathsf{real}} \approx_{(0,\epsilon_{\mathsf{nf}},0)}
    O_{\mathsf{sim}}
    \set{\Phi(m')}.
\]
\end{lemma}

\begin{proof}
The tuple \((0,\epsilon_{\mathsf{nf}},0)\) separates the call into three
parts.
The first and last coordinates are bookkeeping: the decrypt prefix and final
postprocessing are shared deterministic code and therefore spend no KL budget.
The middle coordinate is the only random step whose center changes, namely
the Gaussian sample.
The proof factors the decryption oracle into a shared deterministic prefix and
a noisy continuation.
The prefix reads and increments the decrypt counter, reads the table, checks
that the requested row exists, and branches on whether the two plaintext labels
stored in that row are equal.
Ordinary Hoare reasoning shows that this prefix preserves $\Phi$ and certifies
the table row for the continuation.

Unequal labels are the easy branch.
Both continuations return \texttt{None}, and this branch costs zero.
If the labels are equal, the real continuation samples around
\[
  \mathsf{deterministic\_dec}(sk,c),
\]
while the simulator samples around the table plaintext.
This is exactly the LMSS noise-flooding move: replace the real approximate
decryption center by the ideal table plaintext, and hide the displacement by
adding sufficiently wide discrete Gaussian noise.
The table invariant and perfect correctness give that these centers are within
the ciphertext's recorded error bound.
The chart equation
$d(a,b)=\|I_a(b)\|$ turns this metric statement into a coordinate bound in
$\bbZ^n$.
For each coordinate, the discrete-Gaussian KL calculation gives cost at most
$1/(2\gamma^2)$, hence the vector cost is
$\epsilon_{\mathsf{nf}}=n/(2\gamma^2)$.
The Pythagorean sequencing rules combine the shared prefix, the Gaussian
sampling step, and the shared postprocessing continuation into the displayed
tuple budget.
\end{proof}

\begin{corollary}[Compiled decryption replacement]
\label{cor:compiled-decryption-replacement}
Let \(\Phi_0\) be the initial same-input/same-heap condition for the closed
games.
Then
\[
  \vDash\AEJ\set{\Phi_0}\ 
    H_2 \approx_{\Delta_{\mathsf{comp}}} H_3
    \set{o_L=o_R}.
\]
\end{corollary}

\begin{proof}
Apply \LabTirName{Compile-Replace} to
\Cref{lem:one-call-decryption-replacement}, with \(\mathsf{ODec}\) as the
selected operation.
The trace compiler has already exposed the first \(q\) selected decrypt calls
in \(H_2\) and \(H_3\).
The compiler side conditions are the usual package-validity and memory
restriction obligations: the adversary locations are separated from the
challenger invariant footprint, and the non-decrypt oracle calls preserve
\(\Phi\).
The tuple compiler theorem sums the local budget over the first \(q\) selected
decrypt calls and then applies \LabTirName{Micciancio-Walter}, giving
\[
  \Delta_{\mathsf{comp}}=\sqrt{q\epsilon_{\mathsf{nf}}/2}.
\]
\end{proof}

\subsection{The Exact Bridges}
\label{sec:exact-hops}

The remaining hops are exact, but they are not all the same kind of exactness.
This subsection records what each zero-cost arrow means, so that the single
nonzero loss above is not hiding any informal oracle-replacement step.

The bridge $H_0\to H_1$ is an exact presentation change.
It unfolds the IND-CPAD game from \Cref{app:ind-cpad-game} into open
experiment code linked with the real oracle package.

The bridge $H_1\to H_2$ is ordinary same-package compiler correctness.
The compiler exposes the first $q$ decrypt calls, but those calls are linked
back to the real IND-CPAD oracle package, so the compiled presentation has the
same completed output distribution as the original IND-CPAD game.

The bridge $H_3\to H_4$ is the bounded-query residual-call argument.
After the first $q$ selected decrypt calls have been compiled, the decrypt
counter is already at the bound on every residual decrypt path.
Such calls assert-fail whether they are linked to the real oracle or the
simulator oracle, so the completed output distribution is unchanged.

The bridge $H_4\to H_5$ is ordinary compiler correctness.
Once both compiled and residual decrypt calls are linked to the same simulator
package, removing the compiler wrapper preserves the full completed output,
including heap.

The bridge $H_5\to H_6$ identifies the simulated-decrypt game with the
IND-CPA experiment for $\mathcal{B}_{\cA}$.
Both games sample the same challenge bit and keys, expose the same public keys
to the adversary, maintain the same logical table and decrypt counter, and
answer every adversary oracle call with the same distribution.
The only subtlety is packaging: the direct simulated-decrypt presentation keeps
state in the IND-CPAD challenger heap, while the reduction presentation stores
public challenge state in the outer IND-CPA oracle heap and table/decrypt state
in the reduction heap.
The invariant therefore relates heaps by their adversary-visible table,
counter, public keys, evaluation key, and hidden challenge bit.
Under this relation the final returned predicate is the same Boolean
statement, namely whether the adversary's guess equals the hidden bit.
The Coq proof splits this exact bridge through an auxiliary linked game
$H_6'$ to separate heap-relation reasoning from package-linking correctness;
the paper-level hop is still the single exact identification
$H_5\to\mathsf{IND\text{-}CPA}(\mathcal{B}_{\cA})$.

\begin{theorem}[Noise-flooding security]
\label{thm:noise-flooding-secure}
Let $\epsilon_{\mathsf{nf}}=n/(2\gamma^2)$.
For every IND-CPAD adversary $\cA$ making at most $q$ non-failing decrypt
queries, the verified reduction constructs the IND-CPA adversary
$\mathcal{B}_{\cA}$
such that
\[
  \Pr[H_0(\cA)=1]
  \le
  \Pr[\mathsf{IND\text{-}CPA}(\mathcal{B}_{\cA})=1]
  + 2\sqrt{q\epsilon_{\mathsf{nf}}/2}.
\]
Consequently, using the assumed IND-CPA security bound for the underlying
scheme,
\[
  \Pr[H_0(\cA)=1]
  \le
  \mathsf{security\_bound}_{\mathsf{IND\text{-}CPA}}(\mathcal{B}_{\cA})
  + 2\sqrt{q\epsilon_{\mathsf{nf}}/2}.
\]
\end{theorem}

This is the theorem exposed in Coq as \texttt{NoiseFloodingSecure.is\_secure}.
The intermediate additive-error theorem is the direct formal counterpart of
the hybrid chain above: it relates $H_0$ to $H_6$ by additive
error $\sqrt{q\epsilon_{\mathsf{nf}}/2}$ at the returned-bit postcondition, and
the final theorem packages this into the winning-probability bound.
