\documentclass{article}

\input{packages}

\title{Relational Hoare Logic with Pythagorean Errors}
\author{(Placeholder Authors List)}
\date{\today}

\input{macros}

\begin{document}

\maketitle

\begin{abstract}
    We propose \emph{Relational Hoare Logic with Pythagorean Errors} in order to apply the \emph{Pythagorean probability preservation} property to pairs of probabilistic programs.
    (This is possibly \emph{not} enough material to be a paper draft in and of itself.
    More likely however, this will become part of a larger paper.)
\end{abstract}

\section{Introduction}

In program analysis, pairs of programs are often compared with each other.
For example, we may show that an optimized program implements its specification faithfully, or is equivalent to a reference implementation.
On the other hand, cryptography researchers often prove that certain sub-protocols are indistinguishable from idealized building blocks that are (by definition) composable and secure.
Machinery for these analyses falls under the \emph{relational Hoare Logic} (RHL) framework.
In this work, we present a variant of RHL that considers \emph{Pythagorean probability preservation} of discrete Guassian distributions.

The term \emph{Pythagorean probability preservation} is coined by the work \cite{mw17}, which generalizes an analysis involving Kullback-Leibler divergence in the work \cite{kl-queries}.
Pythagorean probability preservation is used today to argue the security of the ``CKKS" fully homomorphic encryption, as well as to bound the security loss due to floating-point precision limitations in lattice-based cryptography.
Roughly speaking, suppose two distributions $\mathcal{D}$ and $\mathcal{D}'$ are $\varepsilon$-close, and a program $c$ takes $n$ samples from either $\mathcal{D}$ or $\mathcal{D}'$ before outputting $y$ or $y'$ respectively.
By \emph{data processing inequality} or \emph{up-to-bad arguments} then, we would conclude that $y$ and $y'$ are $(n\cdot\varepsilon)$-close.
This standard analysis would however make use of only that $\mathcal{D}$ and $\mathcal{D}'$ are close in total variation distance.
On the other hand, Pythagorean probability preservation achieves an $O(\sqrt{n})$ bound instead \Ethan{What's the scaling on $\varepsilon$?} if $\mathcal{D}$ and $\mathcal{D}'$ are also close in KL-divergence.

The Pythagorean probability preservation property is a cryptographic one;
in other words, it corresponds to analysis at the semantics level.
Much like data-processing inequality is packaged in program logic terms using RHL and up-to-bad arguments then, we present \emph{Relational Hoare Logic with Pythagorean Errors} in order to apply Pythagorean probability preservation directly at the syntactic level.

\section{Preliminaries}

Here we describe the preliminary mathematical and cryptographic background.

\subsection{Notations}

We write $$\Delta(\cD_1, \cD_2)=\frac{1}{2}\sum_x \abs{\cD_1(x)-\cD_2(x)}$$ for statistical distance,
and $$\KL{\cD_1}{\cD_2}=\sum_x \cD_1(x)\ln\parens{\frac{\cD_1(x)}{\cD_2(x)}}$$ for Kullback-Leibler divergence.
For $x\in\bbR^n$, we write $\norm{x}_2=\sqrt{\sum x_i^2}$ for its $\ell_2$-norm.

\subsection{Cryptography Background}

We first recall the formal definition of \emph{Pythagorean Probability Preservation} as given in the work \cite{mw17}.

\begin{definition}
    \label{def:pythagorean-probability-preservation}
    Suppose $\delta(\cD, \cD')$ is a function on two distributions.
    For a parameter $\lambda>0$, we say $\delta$ is $\lambda$-Pythagorean if the following holds:
    
    For all integer $n\in\bbN$ and joint distributions $\set{\cP_i}_{i=1}^n$ and $\set{\cQ_i}_{i=1}^n$ with the same supports $\set{X_i}_{i=1}^n$ such that
    $\delta(\cP_i | a_i, \cQ_i | a_i)\leq \lambda$
    for all $i\in \bbN$ and $a_i\in \set{X_j}_{j=1}^{i-1}$,
    we have
    \[\Delta(\set{\cP_i}, \set{\cQ_i})\leq \norm{\max_{a_i} \delta(\cP_i | a_i, \cQ_i | a_i)}_2.\]
\end{definition}

Restated in terms of Pythagorean probability preservation then,
the work \cite{kl-queries} gives the following theorem:
\begin{theorem}
    \label{thm:kl-is-pythagorean}
    The function $\delta=\sqrt{\dKL}$ is $\lambda$-Pythagorean for all $\lambda\geq 0$.
    \Ethan{Need to double check if there is in fact a constraint on $\lambda$.}
\end{theorem}

\subsection{Discrete Gaussian Distributions}

We denote by $\mathcal{G}_{\mu, \sigma^2}(x)=e^{(x-\mu)^2/\sigma^2}$ the Gaussian function, and its weight as $W_{\mu, \sigma}=\sum_{x=-\infty}^\infty e^{(x-\mu)^2/\sigma^2}$.
We denote the discrete Gaussian distribution as $\mathcal{DG}_{\mu, \sigma^2}$, which satisfies the following probability distribution function:
$$\Pr_{x\gets \mathcal{DG}_{\mu, \sigma^2}}[x=x_0]=\frac{1}{W_{\mu, \sigma^2}}e^{(x_0-\mu)^2/\sigma^2}.$$

Suppose two discrete Gaussian have the same variance and close centers,
then we can then bound the KL-divergence between them.
\begin{lemma}
    For all $\mu, \nu\in\bbZ$ and $\sigma\in\bbR$, we have
	\begin{equation}
		\KL{\DG{\mu}{\sigma^2}}{\DG{\nu}{\sigma^2}} = \frac{\parens{\nu - \mu}^2}{2\sigma^2}
	\end{equation}
\end{lemma}

\section{RHL with Pythagorean Errors}

We now present our logic.
In particular, we reuse the constructions of SSProve \cite{ssprove} where possible, and simply extend the logic in a compatible way to what already exists.

\subsection{Programs and their Semantics}

Our language matches that of the SSProve framework;
for completeness, we sketch its syntax and semantics here.
SSProve encodes cryptographic code using a mix of its computational monad and Rocq's ambient language Gallina.
The monad is defined in \Cref{fig:ssprove-monad}, and an example program is given in \Cref{fig:ssprove-example-program}.
In particular, sequences of sampling operations are encoded using the monadic \texttt{sampler} and \texttt{ret}, much like one would write \texttt{dlet} and \texttt{dunit} in MathComp and EasyCrypt.
Memory accesses are encoded through the monad as well, while pure computations such as control flow are expressed in Gallina.
Finally, to model cryptographic games, the \texttt{opr} instruction allows invoking subroutines throughout the monadic computation.
This can be used to express \emph{memory restrictions} --- for example, to enforce that an adversary cannot simply read the secret key from the internal memory of the challenger.

\begin{figure}
\caption{The Monad}
\label{fig:ssprove-monad}
\begin{minted}{coq}
  Inductive raw_code (A : choiceType) : Type :=
  | ret (x : A)
  | sampler (op : Op) (k : Arit op → raw_code A)
  | getr (l : Location) (k : l → raw_code A)
  | putr (l : Location) (v : l) (k : raw_code A)
  | opr (o : opsig) (x : src o) (k : tgt o → raw_code A).
\end{minted}    
\end{figure}

\begin{figure}
\caption{Example Program}
\label{fig:ssprove-example-program}
\begin{minted}{coq}
#import {sig #[0] : 'nat → 'bool × 'nat } as f ;;
#import {sig #[1] : 'bool → 'unit } as g ;;
'(b,k) ← f 0 ;;
if b then (
  g false ;;
  m ← sample uniform 2 ;;
  ret 0
)
else (
  o ← get ℓ ;;
  #assert (isSome o) as oSome ;;
  let n := getSome o oSome in
  put n := Some (2 + n) ;;
  ret n
)
\end{minted}    
\end{figure}

As with many languages for cryptography, SSProve does not treat unbounded loops.
Finally, as it supports sampling from arbitrary distributions,
no extensions are needed to support discrete Gaussians.
Instead, we simply include the discrete Gaussian $\DG{\mu}{\sigma^2}$ as a possible parameter to the \texttt{sampler} instruction.
For technical details involving semantics of SSProve, we refer the interested reader to the paper \cite{ssprove} and its associated GitHub repository.

\subsection{Judgments and their Validity}

We start with a relational judgment that encodes the right-hand side of \Cref{def:pythagorean-probability-preservation}.
\Ethan{Maybe there is more to be said here.
Finding a ``right" thing to write down was by no means \emph{free}...}

\begin{definition}[Pythagorean Judgment]
Let $\ell>0$ and $\mathcal{E}=(\varepsilon_1,\ldots, \varepsilon_\ell)\in\bbR^\ell$.
Let $\Phi:(X_1\times \cM)\times (X_2\times \cM)\rightarrow\set{0, 1}$ be a predicate on pairs of initial memories and inputs, and $\Psi:Y\times \cM\rightarrow\set{0, 1}$ be a predicate on final memories and outputs.
For two programs $c_1:X_1\rightarrow\code(Y)$ and $c_2:X_2\rightarrow\code(Y)$, we write
\begin{equation}
    \vDash\Pyth \set{\Phi} c_1\approx_{\mathcal{E}} c_2\set{\Psi}
\end{equation}
if for any initial memories $m_1, m_2\in\cM$ and inputs $x_1\in X_1$ and $x_2\in X_2$ satisfying $\Phi(x_1, m_1, x_2, m_2)$,
we have the following:
\begin{itemize}
    \item There exist two distributions $\cP, \cQ\in\distr(\set{Y_i\times M}_{i=1}^\ell)$ where $Y_\ell=Y$ such that:
    \begin{itemize}
        \item $\cP_\ell = \semantic{c_1(x_1)}(m_1)$ and $\cQ_\ell = \semantic{c_2(x_2)}(m_2)$.
        \item For all $t<\ell$ and $a\in \set{Y_i\times M}_{i=1}^{t-1}$,
        let $\cP_{t|a}$ denote $\cP_t$ conditioned on $\cP_{<t}=a$, and similarly for $\cQ_{t|a}$.
        We then have
        $$\KL{\cP_{t|a}}{\cQ_{t|a}} < \varepsilon_t.$$
    \end{itemize}
    \item For all $(y, m')\in \supp\semantic{c_1(x_1)}(m_1)$, we have $\Psi(y, m')$.
\end{itemize}
\end{definition}

We consider also the following total variation distance based relational Hoare judgment which expresses standard up-to-bad arguments.

\newcommand{\UTB}{\mathsf{UTB}}

\begin{definition}[Up-to-bad Judgment]
    Let $\delta > 0$.
    Let $\Phi:(X_1\times \cM)\times (X_2\times \cM)\rightarrow\set{0, 1}$ be a predicate on pairs of initial memories and inputs,
    and let $\Psi:(Y_1\times \cM)\times (Y_2\times \cM)\rightarrow\set{0, 1}$ be a predicate on pairs of final memories and outputs.
    For two programs $c_1:X_1\rightarrow\code(Y_1)$ and $c_2:X_2\rightarrow\code(Y_2)$, we write
    \begin{equation}
    \vDash\UTB \set{\Phi} c_1\approx_{\delta} c_2\set{\Psi}
    \end{equation}
    if for all initial memories and inputs satisfying $\Phi(x_1, m_1, x_2, m_2)$, there exists a coupling $\cP\in\distr(Y_1\times\cM_1\times Y_2\times\cM_2)$ over final memories and outputs such that
    \[\Pr_{(y_1, m_1', y_2, m_2')\gets \cP}[\Psi(y_1, m_1', y_2, m_2')]\geq 1 - \delta.\]
\end{definition}

\subsection{Pythagorean Errors by KL Divergence}

We now present our building blocks of our Pythagorean relational Hoare logic.

Now, with our judgments defined, Pythagorean probability preservation (\Cref{def:pythagorean-probability-preservation} and \Cref{thm:kl-is-pythagorean}) corresponds to the following inference rule:
\[
\inferrule
  {\vDash\Pyth\set{\Phi} \KLs(c_1, c_2) \leq \mathcal{E} \\ \delta = \norm{\mathcal{E}}_2}
  {\vdash c_1\sim_\delta c_2 : m_1=m_2 \implies m_1'=m_2'}
\]
where $m_1, m_2$ are the entire initial memories, and $m_1', m_2'$ are the final memories.    

We now state a few more inference rules, before giving a sketch on how this logic can be used.
To build this KL judgment, we certainly need the ``base cases" as well as sequence rules.
First, two identical programs satisfy this judgment with the right-hand side being the empty tuple $\emptyset$:
\[
\inferrule
  {}
  {\vdash \KLs(c, c) \leq \emptyset}.
\]
On the other hand, if the two programs are both a single randomized assignment instruction, the right-hand side is simply their distance:
\[
\inferrule
  {\forall m, \dKL(\llbracket e_1 \rrbracket(m), \llbracket e_2\rrbracket(m))\leq \varepsilon}
  {\vdash \KLs(v\gets e_1, v\gets e_2) \leq \varepsilon }
\]
where by abuse of notation we write $\varepsilon$ for also the single-element tuple.
The sequence rule then simply concatenates the two lists of divergences, where we write $(\mathcal{E}_1, \mathcal{E}_2)$ as the (flattened) concatenation of the two tuples. 
\[
\inferrule
  {\vdash \KLs(s_1, s_2) \leq \mathcal{E}_1 \\ \vdash \KLs(s_1', s_2') \leq \mathcal{E}_2 \\ c_1=s_1;s_1' \\ c_2=s_2;s_2'}
  {\vdash \KLs(c_1, c_2) \leq (\mathcal{E}_1, \mathcal{E}_2) }
\]

Finally, we sketch how we would use these rules to prove that two programs $c_1$ and $c_2$ are approximately equivalent:
\begin{enumerate}
    \item We first use the sequential composition rule of our KLs judgment in a statement-by-statement way, until we have the KLs judgment on the entirety of $c_1$ and $c_2$.
    \item We then use the Pythagorean rule to turn a KLs judgment into an additive errors judgment.
    \item Finally, we apply \Cref{lem:additive-error-to-pr} to conclude that the outputs of $c_1$ and $c_2$ are indeed close in TVD.
\end{enumerate}

\subsection{RHL with Additive Errors}

As a building block, we start with a judgment that captures typical ``up-to-bad arguments".
We remark that similar judgments exist in many other formulations, which we will discuss in \Cref{sec:related-works}.
For now, we begin with a relational Hoare-style judgment stating that the post-condition is satisfied with high probability.

\begin{definition}
    \label{def:additive-judgment}
    Let $\delta > 0$.
    Let $\Phi$ be a predicate on pairs of initial memories $(m_1,m_2)\in \cM_1\times \cM_2$, 
and let $\Psi$ be a predicate on pairs of final memories.
    We write
    \[c_1\sim_\delta c_2:\Phi\implies \Psi\]
    if for all initial memories $(m_1, m_2)$ satisfying $\Phi$, there exists a coupling $\cP\in\cD(\cM_1\times\cM_2)$ over final memories such that
    \[\Pr_{(m_1', m_2')\gets \cP}[\Psi(m_1', m_2')]\geq 1 - \delta.\]
\end{definition}

As one may expect, this judgment indeed admits many natural inference rules.
For example, one can obtain this judgment by simply sampling from two close distributions:
\[
\inferrule
  {\Delta(\cD_1, \cD_2)\leq \varepsilon \\ v_1, v_2\notin Vars(\Phi)}
  {\vdash v_1\gets\cD_1 \sim_\varepsilon v_2\gets\cD_2:\Phi \Rightarrow (\Phi \wedge v_1 = v_2)}
\]
We show also that this judgment can be used to show that two programs are approximately equivalent.
More concretely, if the post-condition is of the form $y_1=y_2$, then the difference between $\Pr[y_1=y]$ and $\Pr[y_2=y]$ is small:
\begin{lemma}
    \label{lem:additive-error-to-pr}
    Let $\Phi$ be a precondition, and $m_1$ and $m_2$ be initial memories satisfying $\Phi$.
    Suppose we have
    $$c_1\sim_\delta c_2:\Phi\implies(y_1=y_2)$$
    for variables $y_1$ and $y_2$ in the final memory of $c_1(m_1)$ and $c_2(m_2)$ respectively.
    We then have
    \[\abs{\Pr_{m_1'\gets c_1(m_1)}[m_1'(b_1) = \top]-\Pr_{m_2'\gets c_2(m_2)}[m_2'(b_2) = \top]} \leq \delta.\]
\end{lemma}
Finally, a sequential composition rule for this judgment can also be easily written down and proven sound by data processing inequality.
\Ethan{I know I will have to write down a bunch of other boilerplate rules like this and prove their soundness.}

\subsection{Soundness Proofs}

\Ethan{Of course we need to prove that everything is sound.
I will sometime.}

\section{Related Works}
\label{sec:related-works}

We first recall how up-to-bad arguments are typically formalized in RHL and compare it to our additive errors logic.
We then compare our full logic to apRHL \cite{aprhl}.

\subsection{Additive Errors and Up-to-bad Arguments}

Up-to-bad arguments are often formalized using a combination of relational and probabilistic Hoare logic.
More concretely, in RHL we set the post-condition to $\neg B \implies \Psi$, and in pHL we show the probability of $B$ is low.
Frameworks such as EasyCrypt then provide inference rules taking two judgments of the form above and automatically concluding that two programs' outputs are $\varepsilon$-close.
Our additive errors judgment can be seen as simply combining the two judgments above,
with an additional advantage of no longer requiring an explicit description of the ``bad event" $B$.
More precisely, we have
\Ethan{double check that this is true}
\[
\inferrule
  {\vdash c_1 \sim c_2:\Phi\implies(\neg B_2 \Rightarrow\Psi) \\ \vdash \Pr[c_2:\Phi\implies B_2]<\varepsilon}
  {\vdash c_1 \sim_\varepsilon c_2:\Phi\implies\Psi}
\]
where $B_2$ is a predicate on the second program's final memory.

\subsection{Comparison to apRHL}

Approximate relational Hoare logic \cite{aprhl} is another logic for analysis involving KL divergence.
The main motivation is to analyze differential privacy bounds of the form
\[
\Pr[E_1] \leq e^\varepsilon \Pr[E_2] + \delta.
\]
Accordingly, it is based on relational judgments of the form
$c_1\sim_{\langle\varepsilon, \delta\rangle} c_2:\Phi\implies\Psi$.
Roughly speaking, this judgment states that for all initial memories $(m_1, m_2)$, there exists \emph{two} couplings $\cP_L$ and $\cP_R$ such that:
\begin{itemize}
    \item The two couplings are close, with a \emph{multiplicative privacy budget} factor on top of the total variation distance:
    \[
        \sup_E \abs{\Pr_{m_1\gets \cP_L}[E(m_1)] -  e^\varepsilon \Pr_{m_2\gets \cP_R}[E(m_2)]}\leq \delta
    \]
    \item The first marginal of $\cP_L$ matches $\llbracket c_1\rrbracket (m_1)$.
    \item The second marginal of $\cP_R$ matches $\llbracket c_2\rrbracket (m_2)$.
\end{itemize}

Now, first we compare the $\varepsilon = 0$ case of apRHL with our additive errors judgment.
We claim that our judgment is more general than this case of apRHL, in the sense that apRHL judgment implies ours, but not the other way around.
This can also be interpreted as a form of interoperability:
apRHL judgments with $\varepsilon=0$ can be converted to our judgment.

\begin{lemma}
    For all $\Phi$, $\Psi$, $c_1$, $c_2$, $\delta$, suppose we have
    $c_1\sim_{\langle 0, \delta\rangle} c_2:\Phi\implies\Psi$ under apRHL.
    We then have
    $c_1\sim_\delta c_2:\Phi\implies\Psi$ as our additive errors judgment.
    \Ethan{Maybe missing a factor of 2 on one of the $\delta$'s?}
\end{lemma}
\begin{proof}
    \Ethan{Just proof sketch for now.}
    By $\cP_L\approx_\varepsilon \cP_R$, we know that the marginal $\cP_{L, 2}$ is $\varepsilon$-close to $c_2(m_2)$.
    We can therefore define the coupling in our judgment by sampling with weight $1-\varepsilon$ from the ``overlap" of $\cP_{L, 2}$ and $c_2(m_2)$.
    \Ethan{If the previous sentence makes no sense, that's because actually some clever Baysian ``magic" is required.
    I suspect the technique was discovered during our Dilithium project, but never described in the final paper.}
\end{proof}

\begin{lemma}
    There exists $\Phi$, $\Psi$, $c_1$, $c_2$, $\delta$, so that we have $c_1\sim_\delta c_2:\Phi\implies\Psi$ as our additive errors judgment, but $c_1\sim_{\langle 0, \delta\rangle} c_2:\Phi\implies\Psi$ is not true under apRHL.
    \Ethan{Maybe can strengthen to say that for any ``nontrivial" $\delta$ the apRHL judgment is not true with $\delta'=f(\delta)$.}
\end{lemma}
\begin{proof}
    \Ethan{Just proof sketch for now.}
    By construction.
    Take $c_1$ and $c_2$ to both first sample a good/bad bit $b$.
    If $b=0$ then output some stuff that will always satisfy $\Phi$.
    If $b=1$ then output some stuff that will never satisfy $\Psi$.
\end{proof}

Now we compare our full logic with the entirety of apRHL.
Here are the main differences.
\begin{itemize}
    \item Differential privacy allows multiplicative errors in the final security bound, so there is no need for apRHL to remove the $\varepsilon$ factor once it is introduced.
    On the other hand, security of cryptographic primitives (e.g. CKKS) is always additive.
    We therefore do not have a multiplicative factor, and we provide an inference rule to map our Pythagorean judgment back to the additive errors setting.
    \item Pythagorean preservation achieves a sublinear bound in the number of differing sampling operations.
    On the other hand, in apRHL the only inference rules to combine multiple sampling operations cause the errors to grow linearly,
    As a result, any bounds derived using apRHL would be $\Omega(n)$ and no better than by standard data processing inequality.
\end{itemize}

\section*{Acknowledgements}

We thank ChatGPT for insightful technical discussions, and for improving the writing quality of this work.

\bibliographystyle{plain}
\bibliography{reference}

\end{document}
