\section{Preliminaries}

This section fixes the notation and mathematical interface used by the
formalized proof.
The detailed verified analysis appears in later sections; here we record only
the pieces that determine the shape of the security statement.

\subsection{Notation}
\label{sec:prelim-probability-notation}

All distributions in the paper are discrete.
Both security games below sample a uniform challenge bit $b$, receive the
adversary's guess $b^*$, and return the predicate $b^*=b$.  We write
\[
  \WinProb[G]\triangleq \Pr[G=1]=\Pr[b^*=b].
\]
The Rocq development calls this quantity \texttt{winning\_probability}; it
does not separately define or use an advantage normalized around $1/2$.

\begin{definition}[Distance and KL]
Let \(\cD_1\) and \(\cD_2\) be distributions.
Write
\[
  \Delta(\cD_1,\cD_2)
  =
  \frac{1}{2}\sum_x |\cD_1(x)-\cD_2(x)|
\]
for statistical distance, and write
\[
  \KL{\cD_1}{\cD_2}
  =
  \sum_x \cD_1(x)
    \ln\!\left(\frac{\cD_1(x)}{\cD_2(x)}\right)
\]
for KL divergence when the series is well defined.
\end{definition}

\begin{definition}[Summability]
For a real-valued function \(f\) on a discrete type, write
\(\mathsf{summable}(f)\) for MathComp Analysis's summability predicate.
In paper notation, this means that the series \(\sum_x f(x)\) converges in
the sense used by the formal analysis library.
\end{definition}

\begin{definition}[Absolute continuity]
Let \(\cD_1\) and \(\cD_2\) be distributions.
Write
\[
  \mathsf{AbsCont}(\cD_1,\cD_2)
\]
if
\[
  \forall x.\ \cD_2(x)=0 \Rightarrow \cD_1(x)=0 .
\]
\end{definition}

\begin{definition}[Finite KL]
In the Rocq development, well-definedness of KL is tracked by
\(\mathsf{finiteKL}(\cD_1,\cD_2)\), the predicate
\[
  \begin{aligned}
  \mathsf{finiteKL}(\cD_1,\cD_2)
  \quad\Longleftrightarrow\quad&
  \mathsf{AbsCont}(\cD_1,\cD_2)
  \\
  &{}\wedge
  \mathsf{summable}\!\left(
    x\mapsto
    \cD_1(x)\ln\!\left(\frac{\cD_1(x)}{\cD_2(x)}\right)
  \right).
  \end{aligned}
\]
\end{definition}

\begin{definition}[Coupling]
For a joint distribution \(d\) on \(A\times B\), write
\(\mathsf{Cpl}(d;\mu,\nu)\) if \(d\) is a coupling of \(\mu\) and \(\nu\):
its first and second marginals are \(\mu\) and \(\nu\), respectively.
\end{definition}

\begin{definition}[Completed distribution]
For a subdistribution \(\mu\) over a set \(A\), write
\(\completed{\mu}\) for the completed distribution over \(A\cup\{\bot\}\):
\[
  \completed{\mu}(a)=\mu(a)
  \quad(a\in A),
  \qquad
  \completed{\mu}(\bot)=1-\sum_{a\in A}\mu(a).
\]
The point mass at \(\bot\) accounts for the missing mass of \(\mu\).
\end{definition}

For \(x\in\bbR^n\), we write
\(\norm{x}_2=(\sum_i x_i^2)^{1/2}\).
For \(v\in\bbZ^n\), write
\[
  \norm{v}_\infty=\max_i |v_i|,
  \qquad
  d_{\bbZ^n}(u,v)=\norm{u-v}_\infty .
\]
This is the integer-vector distance implemented in the artifact as
\texttt{ivec\_dist}.

We display program-logic rules using the standard inference-rule notation
\[
\inferrule{P_1 \quad P_2}{Q}.
\]
For readers from cryptography, this notation may be read simply as the
implication \(P_1\wedge P_2\Rightarrow Q\): if all premises above the line
hold, then the conclusion below the line follows.
Rules are often displayed with labels, for example
\[
\inferrule*[right=\textsc{Rule-Name}]
  {P_1 \quad P_2}
  {Q}
\qquad\text{or}\qquad
\inferrule[\textsc{Rule-Name}]
  {P_1 \quad P_2}
  {Q}.
\]
The label is the name of the proof principle, much as mathematical lemmas are
named ``triangle inequality'' or ``data processing inequality''; it is not an
additional premise.

\subsection{SSProve Programs and Semantics}
\label{sec:prelim-ssprove-semantics}

SSProve programs are written in the \texttt{raw\_code} monad shown in
\Cref{fig:ssprove-monad}.
The constructors cover pure return, sampling, heap reads and writes, and calls
to imported operations.
As usual in SSProve, cryptographic games are assembled by linking such code
with packages.

\begin{figure}
\caption{SSProve command language used by the formalization}
\label{fig:ssprove-monad}
\begin{minted}{coq}
Inductive raw_code (A : choiceType) : Type :=
| ret (x : A)
| sampler (op : Op) (k : Arit op -> raw_code A)
| getr (l : Location) (k : l -> raw_code A)
| putr (l : Location) (v : l) (k : raw_code A)
| opr (o : opsig) (x : src o) (k : tgt o -> raw_code A).
\end{minted}
\end{figure}

The semantic function is SSProve's \texttt{Pr\_code}.
For an input \(x\) and heap \(m\), a program \(c:X\to\code(Y)\) denotes a
subdistribution
\[
    \semantic{c(x)}(m) \in \distr(Y\times\cM).
\]
Because SSProve semantics may be subprobabilistic, the judgments below use
completed distributions.
For SSProve programs, output/heap pairs are first packed into a uniform choice
type and then completed.
This explicit failure mass is what lets the rules account for failed traces
and other loss of mass without an informal side condition.
No language extension is needed for discrete Gaussians: they are MathComp
distributions embedded into SSProve sampling code by a small sampling wrapper.

When a program is run from input \(x\) and heap \(m\), write
\(\rho=(x,m)\) for the corresponding configuration.
For a program semantics \(\semantic{c(x)}(m)\), we abbreviate its completion
by
\[
  \csemantic{c(x)}(m)\triangleq\completed{\semantic{c(x)}(m)},
\]
and write
\[
  \csemantic{c}_{\rho}\triangleq\csemantic{c(x)}(m).
\]
For programs with output type \(Y\), write
\[
  \Omega_Y=(Y\times\cM)\cup\{\bot\}
\]
for the completed output/heap space, and omit the subscript when the output
type is clear.
If \(\mathcal{T}\) is a distribution over an \(n\)-tuple transcript space,
write \(\mathsf{marg}_n(\mathcal{T})\) for its final-coordinate marginal.
For a prefix \(a\in\Omega^{i-1}\), write \(\mathcal{T}_i\mid a\) for the
conditional distribution of the \(i\)th coordinate given prefix \(a\), using
the conditional-coordinate convention from \Cref{sec:verified-kl-analysis}.
Following the cryptographic literature, we sometimes present such a transcript
law as a correlated family \(\{\mathcal{T}_i\}_{i=1}^n\).
This does not mean the coordinates are independent: formally it is a single
joint law on \(\Omega^n\), with notation \(\mathcal{T}_i\mid a\) for its
conditional coordinate distributions.

\subsection{Metric Charts for Approximate Messages}
\label{sec:prelim-metric-charts}

The noise-flooding theorem is stated for an abstract approximate-FHE scheme.
Its message type \(M\) is equipped with a natural-number-valued distance
\[
  d : M\times M\to\bbN
\]
and an integer-vector chart interface of dimension \(n\).
For each chart center \(a\in M\), the interface provides a map
\[
  I_a : M\to\bbZ^n
\]
and a decoding map
\[
  J_a : \bbZ^n\to M.
\]
The formal proof uses the following chart laws:
\[
  I_a(a)=0,
  \qquad
  d(a,b)=d_{\bbZ^n}(0,I_a(b)),
\]
and, for all centers \(a,b\in M\) and vectors \(x\in\bbZ^n\),
\[
  J_b(x)=J_a\!\left(x+I_a(b)\right).
\]
The first two equations say that every chart is centered at its chosen
plaintext and that the abstract message distance is exactly the integer-vector
distance from the chart origin.
The last equation is the chart-compatibility law used when moving a Gaussian
sample from one plaintext center to another.

This is intentionally the part of the metric structure needed by the security
proof.
For concrete torus-like plaintext spaces, one should read \(I_a(b)\) as the
short representative of \(b-a\) and \(J_a\) as coordinate-wise modular
decoding centered at \(a\).
The usual metric-space laws are expected of concrete instantiations, but the
formal reduction itself only calls on the chart equations above.
\Codex{The artifact does not assume either round-trip equation
$J_a(I_a(b))=b$ or $I_a(J_a(x))=x$.  Thus \texttt{inverse\_isometry} is only a
shift-compatible decoder, not a proved inverse.  State this limitation and the
resulting correctness/utility boundary prominently, or restore the needed
round-trip laws for the intended concrete interpretation.}

\subsection{Pythagorean Probability Preservation}
\label{sec:prelim-pythagorean-preservation}

The quantitative proof uses KL divergence locally and statistical distance
globally.
We recall the parameterized form of Pythagorean probability preservation from
Micciancio and Walter~\cite{mw17}.

\begin{definition}[\(\lambda\)-Pythagorean probability preservation]
\label{def:pythagorean-probability-preservation}
Let \(\delta(\cD,\cD')\) be a nonnegative function on pairs of distributions,
and let \(\lambda>0\).
We say that \(\delta\) is \(\lambda\)-Pythagorean if the following holds.
For any length \(q\), and any two transcript distributions \(\cP,\cQ\) on
\(\prod_{i=1}^q X_i\), suppose that for every coordinate \(i\) and every
prefix \(a\in\prod_{j<i}X_j\),
\[
  \delta(\cP_i\mid a,\cQ_i\mid a)\leq \lambda .
\]
Then
\[
  \Delta(\cP,\cQ)
  \leq
  \left(
    \sum_{i=1}^q
      \left(\sup_a \delta(\cP_i\mid a,\cQ_i\mid a)\right)^2
  \right)^{1/2}.
\]
\end{definition}

The KL-based instance used in this paper applies the definition to the
square-root KL cost
\[
  \delta_{\sqrt{\mathrm{KL}}}(\cD,\cD')
  =
  \sqrt{\frac{D_{\mathrm{KL}}(\cD\Vert\cD')}{2}},
\]
where \(D_{\mathrm{KL}}\) denotes KL divergence.
Equivalently, if we record KL budgets \(s_i\) directly, the same statement
reads as follows:
if every conditional coordinate has finite KL and satisfies
\[
  \KL{\cP_i\mid a}{\cQ_i\mid a}\leq s_i ,
\]
then
\[
  \Delta(\cP,\cQ)
  \leq
  \sqrt{\frac{\sum_i s_i}{2}}.
\]
Thus the formalization is not using a different asymptotic principle; it uses
the budget-tuple specialization of the \(\lambda\)-Pythagorean theorem.
For the uniform noise-flooding case below, each paid decrypt call has
\(s_i=\epsilon_{\mathsf{nf}}\), so this corresponds to
\(\lambda=\sqrt{\epsilon_{\mathsf{nf}}/2}\).
\Cref{sec:verified-kl-analysis} explains the Rocq proof of this KL instance:
first a conditional-coordinate KL chain bound, then Pinsker's inequality.

\subsection{Discrete Gaussian Distributions}

For \(\mu\in\bbZ\) and \(\sigma>0\), define the Gaussian weight
\[
  \mathcal{G}_{\mu,\sigma^2}(x)
  =
  \exp\!\left(-\frac{(x-\mu)^2}{2\sigma^2}\right)
\]
and the normalizing constant
\[
  W_{\mu,\sigma}=\sum_{x\in\bbZ}\mathcal{G}_{\mu,\sigma^2}(x).
\]
The discrete Gaussian \(\DG{\mu}{\sigma^2}\) is the distribution on
\(\bbZ\) given by
\[
  \Pr_{x\gets\DG{\mu}{\sigma^2}}[x=x_0]
  =
  \frac{\mathcal{G}_{\mu,\sigma^2}(x_0)}{W_{\mu,\sigma}}.
\]
The formalized KL calculation used by the noise-flooding proof is
\[
  \KL{\DG{\mu}{\sigma^2}}{\DG{\nu}{\sigma^2}}
  =
  \frac{(\nu-\mu)^2}{2\sigma^2}.
\]
The \(n\)-dimensional noise used in the construction is the product of \(n\)
independent one-dimensional discrete Gaussians, transported through the
integer-vector chart maps.

\subsection{Noise-Flooding Parameters}
\label{sec:prelim-noise-flooding-parameters}

The verified construction takes a positive Gaussian-width multiplier
\(\gamma>0\).
When a ciphertext carries error bound \(B\), the flooded decryption samples an
\(n\)-dimensional discrete Gaussian with standard deviation
\[
  \sigma_B=(B^2+1)\gamma
\]
in each coordinate, then decodes the sampled integer vector through the
plaintext chart.
If two candidate plaintext centers are within distance \(B\), the chart law
above bounds every coordinate displacement by \(B\).
The one-dimensional discrete-Gaussian KL formula therefore gives the
coordinate budget \(1/(2\gamma^2)\), conservatively, and the whole
\(n\)-coordinate decrypt response has per-query KL budget
\[
  \epsilon_{\mathsf{nf}}
  =
  \frac{n}{2\gamma^2}.
\]
For a decryption-query bound \(q\), the Pythagorean preservation theorem gives
compile distance
\[
  \Delta_{\mathsf{comp}}
  =
  \sqrt{\frac{q\epsilon_{\mathsf{nf}}}{2}},
\]
and the final winning-probability theorem pays exactly this distance:
\[
  \mathsf{security\_loss}(q)
  =
  \sqrt{\frac{q\epsilon_{\mathsf{nf}}}{2}}.
\]
The last step uses the tight fact that, for equal-mass distributions, the
difference at any point is at most their statistical distance.  Completed
game-output distributions have mass one, and winning is the Boolean output
point \texttt{true}, so no extra factor is introduced.
These are the parameters appearing in the formal theorem
\texttt{NoiseFloodingSecure.is\_secure}.
