\section{RHL with Pythagorean Errors}
\label{sec:program-logic}

We now describe the program logic used by the formalized noise-flooding proof.
The development does not introduce a new programming language.
Instead, it defines semantic judgments over SSProve \cite{ssprove} programs and proves the inference rules in Rocq.
The central design choice is to keep SSProve's concrete package/linking model, while adding two relational layers: an additive-error judgment for statistical distance arguments and a Pythagorean judgment for conditional KL-divergence bookkeeping.
The probability facts summarized in \Cref{sec:verified-kl-analysis} provide the distribution-level theorems consumed by these rules.

The rules in \Cref{fig:program-logic-core-rules} show the main proof principles used below.
Base rules are proved directly against the SSProve semantics or the distribution library.
Derived rules are still fully verified theorems, but their Rocq proofs mainly compose earlier judgment rules.
The compiler theorems are separated because they are not ordinary structural rules: their soundness rests on an explicit semantic induction over the trace compiler.

\subsection{Judgments and Basic Rules}

We use the SSProve syntax, completed semantics, configuration notation, and
correlated-transcript notation from \Cref{sec:prelim-ssprove-semantics}.
The Rocq formalization represents correlated transcript families directly as
joint laws.

SSProve has exact relational Hoare logic.
For the Pythagorean rules, we use the corresponding one-sided support-style
judgment to supply invariant-preservation facts for shared code, package
calls, and trace residuals.

\begin{definition}[Hoare judgment]
We write
\begin{equation}
\label{eq:hoare-judgment}
    \vDash\mathsf{Hoare}\set{P}\ c\ \set{Q}
\end{equation}
if, for every input $x$ and initial heap $m$ satisfying $P(x,m)$, every output/heap pair in the support of the SSProve semantics of $c(x)$ from $m$ satisfies $Q$:
\[
  \forall x,m.\ 
  P(x,m)\Rightarrow
  \forall y,m'.\
  (y,m')\in\supp\semantic{c(x)}(m)
  \Rightarrow Q(y,m').
\]
\end{definition}

The additive-error judgment is the canonical approximate-equivalence judgment
used for game hops in the paper and in the main Rocq development.
It treats postconditions on successful output/heap pairs as lifted predicates
on completed outputs when needed.
Informally, it says that from any related initial configurations, the two
completed output distributions can be coupled so that the desired relation
fails with probability at most $\varepsilon$.

\begin{definition}[Additive-error judgment]
We write
\begin{equation}
\label{eq:ae-judgment}
    \vDash\AEJ\set{P}\ c_L \approx_\varepsilon c_R\ \set{Q}
\end{equation}
if $0\leq\varepsilon$ and, for every pair of initial configurations
$\rho_L=(x_L,m_L)$ and $\rho_R=(x_R,m_R)$ satisfying $P(\rho_L,\rho_R)$, there exists a coupling $d_{\rho_L,\rho_R}$ of the completed output distributions such that $Q$ fails with probability at most $\varepsilon$:
\[
\begin{aligned}
&\mathsf{Cpl}\!\left(
  d_{\rho_L,\rho_R};
  \csemantic{c_L}_{\rho_L},
  \csemantic{c_R}_{\rho_R}
\right),
\\
&\Pr_{d_{\rho_L,\rho_R}}[Q]\geq 1-\varepsilon .
\end{aligned}
\]
\end{definition}

\Cref{fig:program-logic-core-rules} shows the two main ways to prove an additive-error judgment.
\LabTirName{AE-Couple} constructs the coupling directly, while \LabTirName{AE-TVD-Eq} derives the judgment from a TVD bound on completed outputs.
The formalization also includes consequence, sequencing, and total-output variants.

Compared with EasyCrypt, this is a semantic pRHL/up-to-bad layer with the bad event already internalized into the coupling success probability.
Compared with SSProve out of the box, we reuse SSProve's distributions, couplings, heaps, packages, and linking semantics, but add a reusable approximate judgment over program fragments.

The Pythagorean judgment records a correlated transcript family over completed
outputs rather than merely listing syntactic sampling sites.
Equivalently, and this is how the artifact represents it, it records the
underlying joint transcript distribution and derives the coordinate
conditionals from that joint law.
The transcript is then handed to the probability-level theorem.
The underlying distribution predicate is the Pythagorean distribution predicate
described in \Cref{sec:verified-kl-analysis}.

\begin{definition}[Pythagorean judgment]
Let $s=(s_1,\ldots,s_n)$ be a tuple of real numbers.
We write
\begin{equation}
\label{eq:pyth-judgment}
    \vDash\Pyth\set{P}\ c_L \approx_s c_R\ \set{Q},
\end{equation}
if all entries of $s$ are nonnegative and, for every pair of initial configurations
$\rho_L=(x_L,m_L)$ and $\rho_R=(x_R,m_R)$ satisfying $P(\rho_L,\rho_R)$,
there exist weight-one correlated transcript families
\[
  \{\cP_i(\rho_L,\rho_R)\}_{i=1}^n,
  \{\cQ_i(\rho_L,\rho_R)\}_{i=1}^n
\]
over completed output/heap values such that:
\[
\begin{aligned}
&\cP_n(\rho_L,\rho_R)=\csemantic{c_L}_{\rho_L},
\qquad
\cQ_n(\rho_L,\rho_R)=\csemantic{c_R}_{\rho_R},
\\
&\forall i\in\{1,\ldots,n\},\ 
  \forall a\in\Omega^{i-1}.\
  \mathsf{finiteKL}\!\left(
    \cP_i(\rho_L,\rho_R)\mid a,
    \cQ_i(\rho_L,\rho_R)\mid a
  \right),
\\
&\forall i\in\{1,\ldots,n\},\ 
  \forall a\in\Omega^{i-1}.\
  \KL{\cP_i(\rho_L,\rho_R)\mid a}
     {\cQ_i(\rho_L,\rho_R)\mid a}
  \leq s_i,
\\
&\forall z\in\supp\semantic{c_L(x_L)}(m_L).\ Q(z),
\qquad
\forall z\in\supp\semantic{c_R(x_R)}(m_R).\ Q(z).
\end{aligned}
\]
\end{definition}

\begin{figure}
\centering
\small
\[
\begin{array}{rcll}
J &::=& \vDash\mathsf{Hoare}\set{P}\ c\ \set{Q}
  & \text{support-style unary correctness}\\
  &\mid& \vDash\AEJ\set{P}\ c_L
    \approx_\varepsilon c_R\ \set{Q}
  & \text{additive error}\\
  &\mid& \vDash\Pyth\set{P}\ c_L
    \approx_s c_R\ \set{Q}
  & \text{Pythagorean KL budget}
\end{array}
\]
\begin{mathpar}
\inferrule*[right=\textsc{AE-Couple}]
  {
    0\leq\varepsilon\\
    P(\rho_L,\rho_R)\Rightarrow
      \mathsf{Cpl}(d_{\rho_L,\rho_R};
        \csemantic{c_L}_{\rho_L},
        \csemantic{c_R}_{\rho_R})\\
    P(\rho_L,\rho_R)\Rightarrow
      \Pr_{d_{\rho_L,\rho_R}}[Q]\geq 1-\varepsilon
  }
  {
    \vDash\AEJ\set{P}\ c_L
      \approx_\varepsilon c_R\ \set{Q}
  }

\inferrule*[right=\textsc{AE-TVD-Eq}]
  {
    0\leq\varepsilon\\
    P(\rho_L,\rho_R)\Rightarrow
      \Delta(\csemantic{c_L}_{\rho_L},
             \csemantic{c_R}_{\rho_R})\leq\varepsilon
  }
  {
    \vDash\AEJ\set{P}\ c_L
      \approx_\varepsilon c_R\ \set{o_L=o_R}
  }

\inferrule*[right=\textsc{Pyth-Seq}]
  {
    \vDash\Pyth\set{P}\ c_L\approx_{s_1} c_R\ \set{M}\\
    \vDash\Pyth
      \set{y_L=y_R\wedge m_L=m_R\wedge M(y_L,m_L)}
      k_L\approx_{s_2} k_R\ \set{Q}
  }
  {
    \vDash\Pyth\set{P}
      (y\gets c_L;\ k_L(y))
      \approx_{s_1\mathbin{+\!\!+}s_2}
      (y\gets c_R;\ k_R(y))
      \set{Q}
  }

\inferrule*[right=\textsc{Micciancio-Walter}]
  {
    \vDash\Pyth\set{P}\ c_L\approx_s c_R\ \set{Q}\\
    \delta=\sqrt{\sum_i s_i/2}
  }
  {
    \vDash\AEJ\set{P}\ c_L
      \approx_\delta c_R\ \set{o_L=o_R}
  }
\end{mathpar}
\caption{Core judgment forms and representative rules.  The displayed rules are schematic versions of the Rocq lemmas in \texttt{ProgramLogics/Ae.v} and \texttt{ProgramLogics/Pyth.v}; later rules specialize these principles to discrete Gaussians and compiled oracle calls.}
\label{fig:program-logic-core-rules}
\end{figure}

\subsection{Pythagorean Error Budgets}

The mathematical bridge from conditional KL budgets to TVD is
\Cref{lem:verified-pythagorean-preservation}.
The program logic uses that lemma through a final-marginal TVD theorem: once a Pythagorean judgment has produced transcript distributions satisfying the Pythagorean distribution predicate, the final completed-output marginals are within
\[
  \sqrt{\frac{\sum_i s_i}{2}}
\]
in total variation.
For a Pythagorean budget tuple $s$, we write
\[
  \lVert s\rVert_1 \triangleq \sum_i s_i .
\]
This is the ordinary $\ell_1$ norm because the judgment requires every coordinate of $s$ to be nonnegative.

For the compiler theorem, we use a call-budget tuple $\operatorname{callErr}_q(\varepsilon)$ defined by
\[
\begin{aligned}
\operatorname{callErr}_0(\varepsilon)
  &= (0,\varepsilon,0),\\
\operatorname{callErr}_{q+1}(\varepsilon)
  &= (0,\varepsilon)\mathbin{+\!\!+}
     \operatorname{callErr}_q(\varepsilon).
\end{aligned}
\]
Thus $\operatorname{callErr}_q(\varepsilon)$ contains $q+1$ copies of $\varepsilon$ separated by zero-budget bookkeeping steps, and its TVD bookkeeping lemma is:
\[
    \operatorname{pythTV}
      (\operatorname{callErr}_q(\varepsilon))
    =
    \sqrt{\frac{(q+1)\varepsilon}{2}}.
\]
The oracle-replacement theorem below phrases the bound with a direct call budget $q$, hence $\sqrt{q\varepsilon/2}$.

\subsection{Pythagorean Proof Rules}

\paragraph{Sampling rules.}
The primitive sampling rule embeds an ordinary distribution-level KL fact into a one-coordinate Pythagorean judgment:
\begin{mathpar}
\inferrule[KL-Samp]
  {
    0\leq\varepsilon\\
    \forall x_L,x_R,m_L,m_R.\
      P((x_L,m_L),(x_R,m_R))\Rightarrow m_L=m_R\\
    \forall x_L,x_R.\ \mathsf{finiteKL}(D_L(x_L),D_R(x_R))\\
    \forall x_L.\ dweight(D_L(x_L))=1\\
    \forall x_R.\ dweight(D_R(x_R))=1\\
    \forall x_L,x_R,m_L,m_R.\
      P((x_L,m_L),(x_R,m_R))
      \Rightarrow \KL{D_L(x_L)}{D_R(x_R)}\leq\varepsilon\\
    \forall x_L,x_R,m_L,m_R,y.\
      P((x_L,m_L),(x_R,m_R))\wedge y\in\supp(D_L(x_L))
      \Rightarrow Q(y,m_L)\\
    \forall x_L,x_R,m_L,m_R,y.\
      P((x_L,m_L),(x_R,m_R))\wedge y\in\supp(D_R(x_R))
      \Rightarrow Q(y,m_R)
  }
  {
    \vDash\Pyth\set{P}\
      \mathsf{sample}(D_L)
      \approx_{[\varepsilon]}
      \mathsf{sample}(D_R)
      \set{Q}
  }.
\end{mathpar}
The equal-memory premise is present because sampling returns the sampled value without changing the heap, and the one-step trace encoding packs output/heap pairs.

The discrete Gaussian rule is the corresponding specialization:
\begin{mathpar}
\inferrule[KL-DG]
  {
    0<\sigma\\
    0\leq\varepsilon\\
    \forall \mu_L,\mu_R,m_L,m_R.\
      P((\mu_L,m_L),(\mu_R,m_R))\Rightarrow m_L=m_R\\
    \forall \mu_L,\mu_R,m_L,m_R.\
      P((\mu_L,m_L),(\mu_R,m_R))\Rightarrow
      \frac{(\mu_R-\mu_L)^2}{2\sigma^2}\leq\varepsilon\\
    \forall \mu_L,\mu_R,m_L,m_R,y.\
      P((\mu_L,m_L),(\mu_R,m_R))
      \wedge y\in\supp(\mathcal{DG}_{\mu_L,\sigma^2})
      \Rightarrow Q(y,m_L)\\
    \forall \mu_L,\mu_R,m_L,m_R,y.\
      P((\mu_L,m_L),(\mu_R,m_R))
      \wedge y\in\supp(\mathcal{DG}_{\mu_R,\sigma^2})
      \Rightarrow Q(y,m_R)
  }
  {
    \vDash\Pyth\set{P}\
      x\gets\mathcal{DG}_{\mu_L,\sigma^2}
      \approx_{[\varepsilon]}
      x\gets\mathcal{DG}_{\mu_R,\sigma^2}
      \set{Q}
  }.
\end{mathpar}
The formal rule uses integer centers and the verified KL bound for discrete Gaussians.
There is also a closed-program variant.

\paragraph{Structural rules.}
The reflexivity and consequence rules are proved directly from the transcript semantics:
\begin{mathpar}
\inferrule[Pyth-Refl]
  {
    \forall i.\ 0\leq s_i\\
    \forall x_L,x_R,m_L,m_R.\
      P((x_L,m_L),(x_R,m_R))\Rightarrow x_L=x_R\wedge m_L=m_R\\
    \forall x,m,y,m'.\
      P((x,m),(x,m))\wedge (y,m')\in\supp\semantic{c(x)}(m)
      \Rightarrow Q(y,m')
  }
  {
    \vDash\Pyth\set{P}\ c\approx_s c\ \set{Q}
  }

\inferrule[Pyth-Conseq]
  {
    P'\Rightarrow P\\
    Q\Rightarrow Q'\\
    \forall i.\ s_i\leq s'_i\\
    \vDash\Pyth\set{P}\ c_L\approx_s c_R\ \set{Q}
  }
  {
    \vDash\Pyth\set{P'}\ c_L\approx_{s'} c_R\ \set{Q'}
  }
\end{mathpar}
Notice that \LabTirName{Pyth-Refl} permits any nonnegative budget tuple.
This is just monotonicity of the KL bounds: identical traces have zero KL, so they fit inside any nonnegative budget.

The main sequencing rule concatenates error tuples:
\begin{mathpar}
\inferrule[Pyth-Seq]
  {
    \vDash\Pyth\set{P}\ c_L\approx_{s_1} c_R\ \set{M}\\
    \vDash\Pyth
      \set{y_L=y_R\wedge m_L=m_R\wedge M(y_L,m_L)}
      \ k_L \approx_{s_2} k_R\ \set{Q}
  }
  {
    \vDash\Pyth\set{P}\
      (y\gets c_L;\ k_L(y))
      \approx_{s_1\mathbin{+\!\!+}s_2}
      (y\gets c_R;\ k_R(y))
      \set{Q}
  }.
\end{mathpar}
The second premise is written pointwise: for every intermediate value and heap satisfying $M$, the two continuations are related when run from the same value and heap.
This pointwise formulation avoids a separate syntactic rule for every command constructor.

The mixed Hoare/Pythagorean sequencing rules let the proof run shared code without spending a KL budget.
They are used repeatedly in the compiler proof:
\begin{mathpar}
\inferrule[Pyth-Hoare-Seq]
  {
    P((x_L,m_L),(x_R,m_R))
      \Rightarrow x_L=x_R\wedge m_L=m_R\wedge P_0(x_L,m_L)\\
    \vDash\mathsf{Hoare}\set{P_0}\ c\ \set{M}\\
    \vDash\Pyth
      \set{y_L=y_R\wedge m_L=m_R\wedge M(y_L,m_L)}
      \ k_L\approx_s k_R\ \set{Q}
  }
  {
    \vDash\Pyth\set{P}\
      (y\gets c;\ k_L(y))
      \approx_{[0]\mathbin{+\!\!+}s}
      (y\gets c;\ k_R(y))
      \set{Q}
  }

\inferrule[Pyth-Shared-Cont]
  {
    \vDash\Pyth\set{P}\ c_L\approx_s c_R\ \set{M}\\
    \vDash\mathsf{Hoare}\set{M}\ k\ \set{Q}
  }
  {
    \vDash\Pyth\set{P}\
      (y\gets c_L;\ k(y))
      \approx_{s\mathbin{+\!\!+}[0]}
      (y\gets c_R;\ k(y))
      \set{Q}
  }
\end{mathpar}
There are closed-program variants for hiding unit inputs.

\paragraph{Micciancio-Walter bridge.}
The rule that turns a Pythagorean judgment into an additive-error judgment is:
\begin{mathpar}
\inferrule[Micciancio-Walter]
  {
    \vDash\Pyth\set{P}\ c_L\approx_s c_R\ \set{Q}\\
    \delta=\sqrt{\frac{\sum_i s_i}{2}}
  }
  {
    \vDash\AEJ\set{P}\ c_L
      \approx_\delta c_R\ \set{o_L=o_R}
  }.
\end{mathpar}
The conclusion is intentionally about equality of completed outputs.
The postcondition $Q$ is part of the Pythagorean judgment and is used by sequencing rules, but the bridge theorem extracts the statistical-distance/equality consequence needed for cryptographic indistinguishability.

This is the point where the formalization realizes Pythagorean probability preservation: the final-marginal TVD theorem bounds the TVD of completed outputs, and \LabTirName{AE-TVD-Eq} turns that TVD bound into the additive-error judgment.

\subsection{Trace Compilation and Oracle Replacement}

The Pythagorean rules above explain how to compose a fixed sequence of
KL-budgeted program fragments without losing the $O(\sqrt q)$ bound.
They do not, by themselves, solve the adversarial-oracle problem: an IND-CPAD
adversary decides adaptively when to call the decryption oracle, and the
source program is not presented as a list of $q$ decryption subroutines.
The compiler is the adapter from arbitrary adversary syntax to the fixed-call
schedule expected by the logic.

The oracle-replacement story is therefore not an EasyCrypt-style rule for
opaque adversaries.
Instead, the formalization defines an explicit compilation pass over concrete
SSProve code.
It exposes the first $q$ calls to a selected operation, runs the replacement
proof at those call sites, and replays the residual computation from recorded
traces.
Linked back to the same package, the compiler is semantics-preserving.
Linked to a KL-close replacement package, it gives the quantitative
oracle-replacement theorem used in the noise-flooding proof.

\begin{definition}[Trace]
A trace $\tau$ is a finite list of effect entries
\[
    e ::= \mathsf{call}(y)\mid \mathsf{get}(v)\mid \mathsf{put}\mid \mathsf{sample}(a).
\]
\end{definition}

Intuitively, a trace is the evidence needed to replay the part of the program
that has already been executed.
It records the return value of each non-selected package call, the value read
from each heap location, and the value drawn by each sampling command.
Heap writes are deterministic once the current command is known, so their trace
entry only records that the write occurred.

\begin{definition}[Replay residual]
For a program $c$ and a trace $\tau$, define the partial residual operation
$\operatorname{cont}(c,\tau)$ by replaying $\tau$ against $c$:
\[
\begin{array}{rcl}
\operatorname{cont}(c,[\,])&=&c,\\
\operatorname{cont}(\mathsf{ret}(v),e::\tau)&=&\bot,\\
\operatorname{cont}(\mathsf{opr}(o,x,k),\mathsf{call}(y)::\tau)
  &=&\operatorname{cont}(k(y),\tau),\\
\operatorname{cont}(\mathsf{get}(l,k),\mathsf{get}(v)::\tau)
  &=&\operatorname{cont}(k(v),\tau),\\
\operatorname{cont}(\mathsf{put}(l,v,k),\mathsf{put}::\tau)
  &=&\operatorname{cont}(k,\tau),\\
\operatorname{cont}(\mathsf{sample}(D,k),\mathsf{sample}(a)::\tau)
  &=&\operatorname{cont}(k(a),\tau),
\end{array}
\]
with result $\bot$ when the next trace entry has the wrong tag or payload type.
\end{definition}

The residual operation replays a trace through the original code and returns
the program that remains to be run.
Invalid traces produce $\bot$.
This explicit invalid case lets the Rocq compiler be stated as a total
function, even though intended executions only generate feasible traces.

\begin{definition}[Run to selected call]
Fix an operation identifier $fn$.
Define $\operatorname{runToCall}(c,fn,\tau)$ recursively by executing $c$ until
it either returns or reaches the next call whose identifier is $fn$:
\[
\begin{array}{rcl}
\operatorname{runToCall}(\mathsf{ret}(v),fn,\tau)
  &=& \mathsf{ret}(\mathsf{done}(v),\tau),\\
\operatorname{runToCall}(\mathsf{opr}((fn,S,T),x,k),fn,\tau)
  &=& \mathsf{ret}(\mathsf{query}(x),\tau),\\
\operatorname{runToCall}(\mathsf{opr}(o,x,k),fn,\tau)
  &=& y\gets o(x);\ 
      \operatorname{runToCall}(k(y),fn,\tau\cdot\mathsf{call}(y)),\\
\operatorname{runToCall}(\mathsf{get}(l,k),fn,\tau)
  &=& v\gets \mathsf{get}(l);\ 
      \operatorname{runToCall}(k(v),fn,\tau\cdot\mathsf{get}(v)),\\
\operatorname{runToCall}(\mathsf{put}(l,v,k),fn,\tau)
  &=& \mathsf{put}(l,v);\ 
      \operatorname{runToCall}(k,fn,\tau\cdot\mathsf{put}),\\
\operatorname{runToCall}(\mathsf{sample}(D,k),fn,\tau)
  &=& a\gets D;\ 
      \operatorname{runToCall}(k(a),fn,\tau\cdot\mathsf{sample}(a)).
\end{array}
\]
\end{definition}

The run-to-call pass is the compiler's scanner.
It executes ordinary commands and records them in the trace, but stops just
before the selected operation call.
Thus its result is either $\mathsf{done}(v)$, meaning the program returned, or
$\mathsf{query}(x)$, meaning the next selected call wants input $x$.

\begin{definition}[Factoring selected calls]
Fix an operation identifier $fn$ and a package $P$ implementing that operation.
Define the trace-producing helper
$\operatorname{factorCalls}_q(P,fn,c,\tau)$ by recursion on $q$:
\[
\begin{array}{rcl}
\operatorname{factorCalls}_0(P,fn,c,\tau)
  &=&\mathsf{ret}(\tau),\\
\operatorname{factorCalls}_{q+1}(P,fn,c,\tau)
  &=&
  \begin{cases}
  \mathsf{bad}, & \operatorname{cont}(c,\tau)=\bot,\\
  s\gets\operatorname{runToCall}(c',fn,[\,]);\ F_q(s),
    & \operatorname{cont}(c,\tau)=c',
  \end{cases}
\end{array}
\]
where
\[
\begin{array}{rcl}
F_q(\mathsf{done}(v),\tau')&=&\mathsf{ret}(\tau\cdot\tau'),\\
F_q(\mathsf{query}(x),\tau')
  &=& y\gets P.fn(x);\ 
      \operatorname{factorCalls}_q
        (P,fn,c,\tau\cdot\tau'\cdot\mathsf{call}(y)).
\end{array}
\]
\end{definition}

Factoring repeats the scanner at most $q$ times.
When the scanner finds a selected call, the helper calls \(P.fn\), appends the
observed answer to the trace, and continues scanning the original program from
the resulting residual.
If the program returns before all $q$ calls are found, the current trace is
returned.

\begin{definition}[Trace compiler]
Fix an operation identifier $fn$ and a package $P$ implementing that operation.
The compiled program is
\[
    \operatorname{compileCalls}_q(P,fn,c)
    =
    \tau\gets\operatorname{factorCalls}_q(P,fn,c,[\,]);\
    \begin{cases}
      c',& \operatorname{cont}(c,\tau)=c',\\
      \mathsf{bad},& \operatorname{cont}(c,\tau)=\bot.
    \end{cases}
\]
\end{definition}

The compiler first produces a trace exposing the first $q$ selected calls and
then replays that trace against the source program to obtain the residual code.
This is the point where an adaptive adversary is put into the fixed-call form
expected by the Pythagorean rules.

The formal definitions pack traces and call inputs into SSProve choice types, but the mathematical control flow is the one above.

The same-package correctness theorem is:
\begin{mathpar}
\inferrule[Compile-Same-Package]
  {
    \mathsf{ValidPackage}_{L_P,\emptyset,M}(P)\\
    \mathsf{fhas}(M,fn:X\to Y)\\
    \mathsf{ValidCode}_{L,M}(c)
  }
  {
    \mathsf{codeLink}(
      \operatorname{compileCalls}_q(P,fn,c),P)
    =
    \mathsf{codeLink}(c,P)
  }.
\end{mathpar}
Thus the compiler is not an additional cryptographic assumption; linked with the same package, it is semantics-preserving.

The formalization proves the quantitative replacement theorem for this compiled form.
Write $\mathcal{O}_{P'}(x)$ for \(P'.fn(x)\), and write
$A_{P'}^{(q)}$ for the linked adversary obtained by compiling the first $q$
selected calls through $P'$.
The local assumption is a one-call Pythagorean judgment for the selected operation.
The public replacement rule is stated in its tuple-budget form:
\[
    \vDash\Pyth
      \set{x_L=x_R\wedge m_L=m_R\wedge I(m_L)}
      \mathcal{O}_{P'} \approx_s \mathcal{O}_{P''}
      \set{I(m')}.
\]
The singleton tuple $s=(\varepsilon)$ recovers the scalar one-call rule.
The displayed rules are schematic only in the SSProve packaging parameters.

The remaining side conditions are the SSProve version of what EasyCrypt-style
proofs often call memory restrictions.
They identify the heap region on which the invariant \(I\) is allowed to
depend, and they say that the ordinary package calls made while searching for
the selected call preserve that invariant.

\begin{definition}[Heap agreement on a region]
For heaps \(m_0,m_1\) and a set of locations \(K\), write
\[
m_0\equiv_K m_1
\]
if the heaps agree on every location in \(K\):
\[
\forall \ell.\ \ell\in K\Rightarrow m_0[\ell]=m_1[\ell].
\]
\end{definition}

The set \(K\) is the footprint of the invariant.
If the compiler or adversary changes locations outside \(K\), such changes
should be invisible to \(I\).

\begin{definition}[Invariant footprint]
For a heap predicate \(I\), write
\[
  \mathsf{DependsOnlyOn}_K(I)
\]
if
\[
  \forall m_0,m_1.\ 
  m_0\equiv_K m_1
  \Rightarrow
  (I(m_0)\Leftrightarrow I(m_1)).
\]
\end{definition}

The compiler also executes package calls that are not the selected operation
\(fn\).
Those calls are shared bookkeeping from the point of view of the replacement
proof, so they must preserve the invariant rather than consume Pythagorean
budget.
The selected operation itself is excluded: it is handled by the local
Pythagorean premise.

\begin{definition}[Preservation except at the selected operation]
For an interface \(M\), package \(P'\), operation identifier \(fn\), and heap
predicate \(I\), write \(\mathsf{PresExcept}_{M,P',fn}(I)\) if every operation
in \(M\) other than \(fn\) preserves \(I\):
\[
\begin{aligned}
&\forall o\in M.\ \mathsf{id}(o)\neq fn\Rightarrow\\
&\quad
  \vDash\mathsf{Hoare}
    \set{(u,m)\mapsto I(m)}
\\
&\quad\quad
    (u\mapsto P'.o(u))
\\
&\quad\quad
    \set{(v,m')\mapsto I(m')}.
\end{aligned}
\]
\end{definition}

\begin{definition}[Compiler side conditions]
Let $\mathcal{H}_{\mathsf{cmp}}$ abbreviate the following conjunction:
\[
\begin{aligned}
\mathcal{H}_{\mathsf{cmp}}
\Longleftrightarrow{}&
(\forall x.\ \mathsf{ValidCode}_{L,M}(\mathsf{prog}(x)))\\
&{}\wedge
\mathsf{ValidPackage}_{L',\emptyset,M}(P')
\wedge
\mathsf{ValidPackage}_{L'',\emptyset,M}(P'')\\
&{}\wedge
\mathsf{fseparate}(K,L)
\wedge
\mathsf{fhas}(M,fn:X\to Y)\\
&{}\wedge
\mathsf{DependsOnlyOn}_K(I)
\wedge
\mathsf{PresExcept}_{M,P',fn}(I).
\end{aligned}
\]
\end{definition}

Thus \(\mathcal{H}_{\mathsf{cmp}}\) collects the well-formedness obligations
for the source code and packages, the disjointness condition saying that the
compiled code's locations \(L\) are separate from the invariant footprint
\(K\), the presence of the selected operation \(fn\), and the two memory
restriction predicates above.
The fully quantified theorem keeps the SSProve location, interface, and package-validity parameters explicit.

Let \(P_I\) abbreviate the synchronized initial condition
\[
  x_L=x_R\wedge m_L=m_R\wedge I(m_L).
\]
For readability, the intermediate Pythagorean compiler rule is shown as a
scalar instance of the tuple-budget theorem used by
\LabTirName{Compile-Replace}.
\begin{mathpar}
\inferrule[Pyth-Compile-Calls]
  {
    \mathcal{H}_{\mathsf{cmp}}\\
    \vDash\Pyth_1\set{P_I}\ \mathcal{O}_{P'}
      \approx_\varepsilon \mathcal{O}_{P''}\ \set{I(m')}
  }
  {
    \vDash\Pyth\set{P_I}
      \ A_{P'}^{(q+1)}
      \approx_{\operatorname{callErr}_q(\varepsilon)}
      \ A_{P''}^{(q+1)}
      \set{I(m')}
  }.
\end{mathpar}
The proof is by induction over the call bound.
Each inductive step spends one $\varepsilon$ entry for the selected package
call.
Hoare rules handle the shared search/resume code around it.
This compiler family is the most involved program-logic proof in the
development: it combines the call-bound induction with residual/replay
correctness, invalid-trace cases, heap-region preservation, and tuple-budget
bookkeeping.

This is the oracle-replacement statement needed by the noise-flooding proof.
Informally, if one selected oracle implementation is Pythagorean-close to
another with tuple budget \(s\), then replacing the first \(q\) selected calls
inside a compiled adversary changes the final experiment by at most
\[
    O_i\sim_s O_i'
    \quad\Longrightarrow\quad
    A^{O_1,\ldots,O_i,\ldots}
    \sim_{\sqrt{q\lVert s\rVert_1/2}}
    A^{O_1,\ldots,O_i',\ldots}.
\]
The tuple form is the one used below: the selected decryption body is factored
into a shared deterministic prefix, one discrete-Gaussian sampling coordinate,
and a shared deterministic continuation.

Composing the compiler theorem with \LabTirName{Micciancio-Walter} gives the
formal rule:
\begin{mathpar}
\inferrule[Compile-Replace]
  {
    \mathcal{H}_{\mathsf{cmp}}\\
    \vDash\Pyth\set{P_I}\ \mathcal{O}_{P'}
      \approx_s \mathcal{O}_{P''}\ \set{I(m')}
  }
  {
    \vDash\AEJ
      \set{P_I}
      \ A_{P'}^{(q)}
      \approx_{\sqrt{q\lVert s\rVert_1/2}}
      \ A_{P''}^{(q)}
      \set{o_L=o_R}
  }.
\end{mathpar}

Compared with EasyCrypt, the difference is mostly where the abstraction boundary sits.
EasyCrypt's adversarial call rules reason about opaque adversary modules.
Here the adversary is executable syntax, so the proof factors the adversary by a verified trace compiler.
Compared with stock SSProve, the same-package theorem is ordinary package-linking correctness, while \LabTirName{Compile-Replace} adds the quantitative statement that replacing one selected operation by a KL-close implementation changes the compiled adversary's completed output distribution by at most $\sqrt{q\lVert s\rVert_1/2}$.

\input{related-works}
