\section{Comparison with Other Program Logics}
\label{sec:related-works}

This section situates the verified logic against nearby relational program
logics.
The short version is that our additive-error layer is close in spirit to
standard up-to-bad reasoning, while the Pythagorean layer and trace compiler
are the pieces that are specific to the noise-flooding proof.
They let us package a one-call KL argument as a whole-game theorem with
loss $\sqrt{q\lVert s\rVert_1/2}$ for a concrete SSProve adversary
making up to $q$ selected oracle calls.

\subsection{Additive Error and Up-to-bad Reasoning}
\label{sec:up-to-bad-event}

The most familiar cryptographic comparison principle is the up-to-bad
argument.
One proves a relational judgment saying that two executions agree unless a bad
event occurs, and separately proves that the bad event has small probability.
In an EasyCrypt-style presentation, this often appears as an RHL judgment with
postcondition $\neg B\Rightarrow\Psi$, followed by a probabilistic Hoare proof
that $\Pr[B]\leq\varepsilon$.
The conclusion is a statistical-distance or distinguishing-advantage bound.

Our additive-error judgment internalizes this final step.
The judgment
\[
  \vDash\AEJ\set{P}\ c_L
    \approx_\varepsilon c_R\ \set{Q}
\]
directly asserts a coupling of the completed output distributions satisfying
$Q$ with probability at least $1-\varepsilon$.
Thus the bad event need not be named when a direct coupling or TVD argument is
more convenient.
Conversely, the usual up-to-bad pattern embeds into our judgment: if a coupling
satisfies $\neg B\Rightarrow Q$ and the coupled probability of $B$ is at most
$\varepsilon$, then the same coupling witnesses the additive-error judgment
with loss $\varepsilon$.

This is the role of the additive-error layer in the noise-flooding proof.
It is not where the sublinear bound is created.
It provides the completed-output notion of closeness, exact bridge rules, and
triangle rules used to compose the hybrid chain in \Cref{sec:games}.

\subsection{Where Sublinear Error Enters}

Standard relational logics compose approximate judgments linearly.
If a sequence rule is applied to two fragments with losses $\varepsilon$ and
$\varepsilon'$, the composed fragment typically has loss
$\varepsilon+\varepsilon'$.
This is the right shape for union-bound and up-to-bad proofs, but it does not
capture the Micciancio-Walter Pythagorean phenomenon used in lattice
cryptography.

Our Pythagorean judgment records a tuple of conditional KL costs rather than an
additive TVD cost.
Only after the relevant transcript has been built do we apply the verified
Pythagorean preservation theorem, yielding the final-marginal bound
\[
  \Delta \leq \sqrt{\frac{\sum_i s_i}{2}}.
\]
For the oracle setting, \LabTirName{Compile-Replace} combines this with the
trace compiler: from a one-call Pythagorean judgment for a selected oracle
operation, it derives a additive-error judgment for the whole
compiled adversary, with loss
\[
  \sqrt{q\lVert s\rVert_1/2}.
\]
This is the proof boundary used by the paid hop $H_2\to H_3$ in
\Cref{sec:paid-hop}.

\subsection{Concrete Adversaries and Trace Compilation}

EasyCrypt-style developments often reason about adversaries through module
interfaces and adversarial call rules.
This is a good abstraction for cryptographic proofs, and reflection techniques
can make distribution-level proof steps available at the program level
\cite{reflection}.
SSProve \cite{ssprove} has a different foundation: adversaries and games are
concrete packages whose semantics are defined in Coq.

The trace compiler is our bridge between these worlds.
Instead of assuming that an adversary is already written as $q$ decryption
subroutines, the compiler searches the concrete program for the first $q$
calls to a selected operation identifier, records the surrounding trace, and
resumes the residual program.
The theorem \LabTirName{Compile-Same-Package} says that compiling and then
linking back to the same package is semantics-preserving.
The theorem \LabTirName{Compile-Replace} says that replacing those selected
calls by a Pythagorean-close package changes the completed output distribution
by the Pythagorean loss above.

This trace-compiler proof is one of the main proof-engineering contributions
of the development.
It accounts for invalid replay traces, heap-region invariants, residual
operation calls, and the fact that an ordinary adversary may decide when to
query decryption in a data-dependent way.

\subsection{Comparison with apRHL}

Approximate relational Hoare logic and its approximate-coupling extensions
were designed primarily for differential privacy
\cite{aprhl-couplings}.
The related approximate-lifting framework for $f$-divergences gives a broader
semantic account of the same style of reasoning \cite{aprhl}.
The characteristic judgment has parameters $(\varepsilon,\delta)$ and supports
privacy bounds of the form
\[
  \Pr[E_1] \leq e^\varepsilon \Pr[E_2]+\delta .
\]
This multiplicative component is essential in differential privacy, but it is
not the usual shape of a game-based cryptographic advantage bound.
If one wants an ordinary additive distinguishing bound, the multiplicative
term must be converted away, for instance by the coarse estimate
\[
  \Pr[E_1]-\Pr[E_2]
    \leq (e^\varepsilon-1)\Pr[E_2]+\delta
    \leq e^\varepsilon-1+\delta,
\]
and a symmetric bound requires a corresponding judgment in the other
direction.
Moreover, the standard sequential composition of approximate liftings adds
the privacy parameters linearly.
This is exactly the behavior our Pythagorean layer is meant to avoid for
KL-based cryptographic analyses.

The comparison is cleanest at $\varepsilon=0$.
For clarity, we recall the relevant semantic shape of the apRHL judgment.

\begin{definition}[apRHL judgment]
  The judgment
  \[
    \vDash\apRHL\set{P}\ c_L
      \sim_{\langle\varepsilon,\delta\rangle} c_R\ \set{Q}
  \]
  means that for every pair of initial states satisfying $P$, the output
  distributions $\mu_L$ and $\mu_R$ of $c_L$ and $c_R$ admit an
  $(\varepsilon,\delta)$-approximate lifting of $Q$.
  Concretely, this means there are witness distributions
  $\eta_L,\eta_R$ over pairs of outputs such that the first marginal of
  $\eta_L$ is $\mu_L$, the second marginal of $\eta_R$ is $\mu_R$, both
  witnesses are supported in $Q$, and
  \[
    \Delta_\varepsilon(\eta_L,\eta_R)
      \triangleq
      \sup_E
        \left(\Pr_{\eta_L}[E] - e^\varepsilon\Pr_{\eta_R}[E]\right)
      \leq \delta .
  \]
  At $\varepsilon=0$, for full output distributions, this divergence is the
  usual total-variation distance between the two witnesses.
\end{definition}

The structural difference from our additive-error judgment is important.
An apRHL proof uses two coupling-like witnesses, both of which must satisfy
the postcondition exactly; the error is charged to the distance between those
two witnesses.
Our additive-error proof uses one coupling of the actual output distributions;
the error is charged directly to the probability that the postcondition fails.
The next two lemmas make this distinction precise at $\varepsilon=0$.

\begin{lemma}[Zero-privacy apRHL implies additive error]
  If
  \[
    \vDash\apRHL\set{P}\ c_L
      \sim_{\langle 0,\delta\rangle} c_R\ \set{Q},
  \]
  then
  \[
    \vDash\AEJ\set{P}\ c_L
      \approx_\delta c_R\ \set{Q}.
  \]
\end{lemma}

\begin{proof}
  Fix related initial states and let $\eta_L$ and $\eta_R$ be the apRHL
  witnesses for the resulting output distributions.
  Since $\varepsilon=0$, the definition gives
  $\TVD(\eta_L,\eta_R)\leq\delta$.
  Let $\kappa$ be a maximal coupling of $\eta_L$ and $\eta_R$.
  Thus
  \[
    \Pr_{(z_L,z_R)\gets\kappa}[z_L=z_R]
      = 1-\TVD(\eta_L,\eta_R)
      \geq 1-\delta .
  \]
  Write $z_L=(x_L,y_L)$ and $z_R=(x_R,y_R)$.
  Define a distribution $\gamma$ over output pairs by sampling
  $(z_L,z_R)\gets\kappa$ and returning $(x_L,y_R)$.
  The left marginal of $\gamma$ is the first marginal of $\eta_L$, hence the
  output distribution of $c_L$; the right marginal of $\gamma$ is the second
  marginal of $\eta_R$, hence the output distribution of $c_R$.
  Therefore $\gamma$ is a coupling of the two program outputs.
  Moreover, whenever $z_L=z_R$, the returned pair is exactly that common
  witness pair, which lies in $Q$ because both witnesses are supported in $Q$.
  Hence $\Pr_\gamma[Q]\geq 1-\delta$, as required by
  $\AEJ$.
\end{proof}

Thus the old intuition that ``apRHL at $\varepsilon=0$ implies additive
error'' is correct, and no extra factor of two is needed for full
distributions.

\begin{lemma}[A same-postcondition separation]
  Let $0<\delta\leq 1$, let
  $c_\delta$ be the program that samples $x\gets\Bern(1-\delta)$ and returns
  $x$, and let $Q(x_L,x_R)$ mean $x_L=x_R=\mathsf{true}$.
  Then
  \[
    \vDash\AEJ\set{\top}\ c_\delta
      \approx_\delta c_\delta\ \set{Q},
  \]
  but for every $\delta'$,
  \[
    \not\vDash\apRHL\set{\top}\ c_\delta
      \sim_{\langle 0,\delta'\rangle} c_\delta\ \set{Q}.
  \]
\end{lemma}

\begin{proof}
  Let $\mu$ be the distribution $\Bern(1-\delta)$.
  For the additive-error judgment, use the diagonal coupling of $\mu$ with
  itself: sample $b\gets\mu$ and return $(b,b)$.
  This is a coupling of the two output distributions of $c_\delta$.
  It satisfies $Q$ exactly when $b=\mathsf{true}$, which happens with
  probability $1-\delta$.
  Therefore the additive-error judgment holds with loss $\delta$.

  For the apRHL judgment, suppose toward contradiction that the judgment held
  for some $\delta'$.
  Let $\eta_L$ and $\eta_R$ be its witnesses.
  Since both witnesses must be supported in $Q$, they are supported only on
  the single pair $(\mathsf{true},\mathsf{true})$.
  The first marginal of $\eta_L$ is therefore the point mass at
  $\mathsf{true}$.
  But the first marginal is required to be the output distribution $\mu$ of
  $c_\delta$, and $\mu(\mathsf{false})=\delta>0$.
  This contradiction is independent of $\delta'$, so no such apRHL judgment
  exists.
\end{proof}

This separation should not be read as a broad completeness claim against
apRHL.
One can often recover a traditional proof by adding an explicit bad event to
the postcondition and proving it separately.
Rather, the point is that our additive-error judgment puts the probability of
postcondition failure directly in the judgment, which is the form needed by
the completed-output hybrid proofs in this paper.
